---
title: "RagePackageTest"
author: "Sophie Wulfing"
date: "2023-02-23"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev="cairo_pdf")
#knitr::write_bib look into
setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Mtx_Model")

library(tinytex)
library(matlib)
library(ggplot2)
library(quadprog) #cite
library(gridtext)
library(popbio) #cite
library(knitr)
library(gridExtra)
library(RColorBrewer) 
#library(rPref) #Double check what this is doing. Maybe cite?
library(tidyr)
library(dplyr)
library(ggpubr)
library(bookdown)
library(kableExtra)


# #Raw data n supplemental. ASK
# in-stage bar graph goes in paper
# in stage management scenarios will definitely be supplemental
# Check out ecological modelling journal
# Write about stable stage cause I don't think you do
# Citations! Also check ahilya table
# APPENDIX!
# MAKE FIG CAPS FOR EVERYTHING AND EDIT FIGS.



observedData <- as.matrix(data.frame(read.csv("RaberinaryBenbow2012Calcs.csv")))
observedStageMatrix <- observedData[1:13, 2:5]
```
```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
```{r FXNoverallsurvive}
survivability <- function(increase){
    #Calculate survivability for each staege and increase by x%
    NewMtx <- (A * increase) + A
    
    for(i in 1:ncol(NewMtx)-1){
      if(sum(NewMtx[,i]) > 1 | NewMtx[4,4] > 1){
        return(NA)
      }}
    return(NewMtx)
}


#This is stupid but it just forces R to stop writing things as complex numbers
f <- function(x) {
   if (all(Im(z <- zapsmall(x))==0)) as.numeric(z) else x
}

real_check <- function(mtx){
  if(sum(mtx[,1]) > 1 | sum(mtx[,2]) > 1 | sum(mtx[,3]) > 1 | mtx[4,4] > 1){
    return(NA)
  }
  else{return(mtx)}
}

```
```{r QuadProg, echo = FALSE}
#Make Z
z <- matrix(t(observedStageMatrix[2:11,]), ncol = 1, nrow = 40, byrow = TRUE)
#z

#Make M
m <- matrix(, nrow = nrow(z) , ncol = 8)
for(i in 1:(nrow(observedStageMatrix)-3)){
  m[(4 * (i-1) + 1):(4 * i),] <- matrix(c(observedStageMatrix[i,1], 0, 0, 0, 0, 0, observedStageMatrix[i,4], 0,
                                          0, observedStageMatrix[i,1], observedStageMatrix[i,2], 0, 0, 0, 0, 0,
                                          0, 0, 0, observedStageMatrix[i,2], observedStageMatrix[i,3], 0, 0, 0,
                                          0, 0, 0, 0, 0, observedStageMatrix[i,3], 0, observedStageMatrix[i,4]),
                                        nrow = 4, byrow = TRUE)
  #m
}

#Make C
c <- matrix(c(diag(8) * (-1), 
              1, 1, 0, 0, 0, 0, 0, 0,
              0, 0, 1, 1, 0, 0, 0, 0,
              0, 0, 0, 0, 1, 1, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 1), nrow = 12, ncol = 8, byrow = TRUE)
#c
#Make b
b <- matrix(c(rep(0,8), rep(1,4)), nrow = 12, ncol = 1, byrow = TRUE)

#Make P
#p <- matrix(, nrow = 8, ncol = 1, byrow = TRUE) #vector(p1, g1, p2, g2, p3, g3, f4, p4)

#Make mtx G and vector x
g <- t(m) %*% m
f_t <- -(t(-z) %*% m)

qp <- solve.QP(Dmat = g, dvec = f_t, Amat = -t(c), bvec = -t(b))#, factorized = FALSE)


A <- matrix(c(qp$solution[1], 0, 0, qp$solution[7],
                 qp$solution[2], qp$solution[3], 0, 0,
                 0, qp$solution[4], qp$solution[5], 0,
                 0, 0, qp$solution[6], qp$solution[8]), byrow = TRUE, nrow = 4, ncol = 4)
#A
```
```{r iteration}
#Calculates new rates based on vectors of % chagne and frequencies of restrictions
percentage <-rates <- c(seq(0, 1, .01)) # c(0.01, 0.05, 0.1, 0.15)
freqs <- c(seq(1, 12)) #c(1, 2, 3, 4, 6, 12)
iteration <- data.frame(matrix(ncol=4,nrow=0, dimnames=list(NULL, c("Frequency", "PercentChange", "mu", "lambda"))))

for(i in 1:length(freqs)){
  for(j in 1:length(percentage)){
    if (!is.na(survivability(percentage[j]))){
      mu <- ((freqs[i]-1) * log(eigen(A)$values[1]) + log(eigen(survivability(percentage[j]))$values[1]))/ freqs[i]
      lambda <- exp(mu)
          iteration[nrow(iteration) + 1,] <- c(freqs[i], percentage[j], f(mu), f(lambda))
    }
      # neweig <- (eigen(survivability(percentage[j]))$values[1] + (freqs[i]-1) * eigen(A)$values[1]) / freqs[i]
  }
}
#kable(iteration)

# Sources: https://www.webpages.uidaho.edu/wlf448/stochgro.pdf
# https://esajournals.onlinelibrary.wiley.com/doi/10.2307/1943004
```

# Multi Month Closures

```{r multiMoClosures, fig.cap = "Projections of different lengths of closures. Agrees with the 7% increase needed for 3 month closures. Double check with other graph."}

####Note, this is calculated differently than the idk random block as this was calculated by multiplying the matrices, not the eigenvalues
p <- list()
modeled_data <- data.frame(matrix(ncol=8,nrow=0, dimnames=list(NULL, c("Stage_1", "Stage_2", "Stage_3", "Stage_4", "month", "pop", "Frequency", "PercentChange"))))
percentage <-rates <- c(seq(0, .1, .01)) # c(0.01, 0.05, 0.1, 0.15) 
freqs <- c(seq(1, 12)) #c(1, 2, 3, 4, 6, 12) 
#par(mfrow = c(6,4))
for(i in 1:length(freqs)){
  for(j in 1:length(percentage)){
    pop <- observedStageMatrix[13,]
    N <- list()
    for (k in 1:120){
      if(k %% 12 %in% c(seq(1, freqs[i]))){
      N[[1]] <- pop
      N[[k+1]] <- survivability(percentage[j]) %*% N[[k]]
      }
    else{
       N[[1]] <- pop
      N[[k+1]] <- A %*% N[[k]]
      }}
   
  
  new_data <- t(as.data.frame(do.call(cbind, N)))
  colnames(new_data) <- c("Stage_1", "Stage_2", "Stage_3", "Stage_4")
  new_data <- data.frame(cbind(new_data, month = 0:120, pop = rowSums(new_data[,1:4]), Frequency = freqs[i], PercentChange = percentage[j]))
  modeled_data <- rbind(modeled_data, new_data)
  }} 

#### CALCULATE GROWTH RATE BASED ON LAST POP-FIRST POP/FIRST POP FOR EACH FREQ/PERCENT COMBINATION
growthRate <- modeled_data %>%
  select(month, pop, Frequency, PercentChange) %>%
  filter(month == 0 | month == 120) %>%
  pivot_wider(names_from = month, values_from = pop) %>%
  rename(month_0 = 3,
         month_120 = 4) %>%
  mutate(Mu_New = log(month_120/month_0)/120) %>% #See Crouse paper. From Nt = N0e^(rt)
  mutate(lambda_new = exp(Mu_New)) %>%
  select(Frequency, PercentChange, Mu_New, lambda_new)

  

#### COPY PARETTO GRAPH FROM OTHER ANALYSIS
paretospagheto <- growthRate %>% 
  filter(lambda_new >= 1) %>%
  group_by(Frequency) %>%
  filter(lambda_new == min(lambda_new))

paretospagheto$Frequency <- paretospagheto$Frequency - 0.5
paretospagheto$PercentChange <- paretospagheto$PercentChange - 0.005


#Continuous Colors
ggplot(growthRate, aes(Frequency, PercentChange, fill = lambda_new)) + geom_tile(width=1) +
  scale_fill_gradient2(
  low = "red",
  mid = "white",
  high = "green",
  #limits = c(1, max(growthRate$lambda_new)),
  midpoint = 1,
  space = "Lab",
  na.value = "grey50") +
  scale_x_continuous(name = "Length of closure (months)", expand=c(0,0), breaks = c(1, 3, 6, 9, 12)) +
  scale_y_continuous(name = "Percent change in survival", expand=c(0,0)) +
  labs(fill = "Eigenvalue") +
  geom_segment(aes(x = 2.5, y = .075, xend = 2.5, yend = 0.105), size = .5, color = "#5A5A5A") +
  geom_segment(aes(x = 11.5, y = .015, xend = 12.5, yend = 0.015), size = .5, color = "#5A5A5A") +
  geom_step(data = paretospagheto , aes(x=Frequency, y=PercentChange), size = .5, color = "#5A5A5A") 

```

\newpage

# Rage Package info

This package also has the following functions I didn't bother with:

* mpm_collapse which collapses a matrix population model to a smaller number of stages

* mpm_rearrange which rearranges stages of a matrix population model to segregate reproductive and non-reproductive stages

* name_stages allows user to name stages

* perturb_stochastic calculates stochastic elasticities from a time-series of matrix population models and corresponding population vectors

* pop_vectors derives a hypothetical set of population vectors corresponding to a
time-series of matrix population models

* repro_stages Identify which stages in a matrix population model are reproductive

* shape_rep Calculate shape of reproduction over age. TBH I just didn't get this one

* shape_surv Calculate shape of survival over age. Same as above

* standard_stages Identify stages corresponding to different parts of the reproductive life
cycle



Note: Not using the default matU or matR. Double check that

```{r ragepackage, echo = FALSE}
library(Rage)
matU <- mpm_split(A)$matU
matU[1,1] <- A[1,1]

matR <- mpm_split(A)$matF
matR[1,1] <- 0

#mpm_standardize is the opposite of mpm_split

```

## Age from Stage

Age specific calculations from Rage package

```{r agefromstage, results = 'asis', echo = FALSE}

#make mtx with 1_13 rows and each col is vector below
agefromstage <- data.frame(matrix(, nrow = 13, ncol = 5))
colnames(agefromstage) <- c("Age", "Reproduction", "Survivorship", "SurvivalProb", "MortalityHazard")

agefromstage$Age <- c(seq(1, 13))

#Age Specific reproduction
mx <- c(mpm_to_mx(matU, matR, start = 1L, xmax = 1000, lx_crit = 0.01, tol = 1e-04))
#Age Specific Survivorship
lx <- c(mpm_to_lx(matU, start = 1L, xmax = 1000, lx_crit = 0.01, tol = 1e-04))
#Age Specific Survival probability
px <- c(mpm_to_px(matU, start = 1L, xmax = 1000, lx_crit = 0.01, tol = 1e-04))
#Age Specific Mortality Hazard
hx <- c(mpm_to_hx(matU, start = 1L, xmax = 1000, lx_crit = 0.01, tol = 1e-04))


agefromstage$Reproduction <- mx
agefromstage$Survivorship <- lx
agefromstage$SurvivalProb <- px
agefromstage$MortalityHazard <- hx

  knitr::kable(agefromstage, format = "latex", booktabs = TRUE) #%>%
  # column_spec(2:3, width = "4.5cm") %>%
  # kable_styling(latex_options="scale_down")
```

\newpage

## Life Expectancy and longevity

Life expectancy applies Markov chain approaches to calculate. Longevity - (the age x at which survivorship for a synthetic cohort falls below some critical proportion (in this case 0.01). 

Note: Life expectancy and longevity are very different. Wut?

```{r lifeExp, results = 'asis', echo = FALSE}
lem <- life_expect_mean(matU, start = 1L)
lev <- life_expect_var(matU, start = 1L)
#le <- life_expect(matU, start = 1L)
lon <- longevity(matU, start = 1L, x_max = 1000, lx_crit = 0.01)

lifeexptable <- data.frame(matrix(, ncol = 2, nrow = 3) )

lifeexptable[1] <-  c("Life Expectancy", "Life Exp Variance", "longevity")
lifeexptable[2] <- c(lem, lev, lon)

knitr::kable(lifeexptable, format = "latex", booktabs = TRUE)

```

\newpage

## Life Table

* x age at the start of the age interval [x, x+1)
* Nx The number of individuals alive at age x. The initial number is set with radix
* Dx proportion of original cohort dying during the age interval [x, x+1)
* lx survivorship, defined as the proportion of initial cohort surviving to the start of age interval [x, x+1)
* dx proportion of original cohort dying in the age interval [x, x+1)
* ax The average time survived within the interval by those that die during the age interval [x, x+1). Assumed to be 0.5
* hx force of mortality (hazard) during the age interval [x, x+1)
* qx probability of death during the interval [x, x+1) for those entering the interval
* px probability of survival for the interval [x, x+1) for those entering the interval
* Lx total person-years lived during the interval [x, x+1)
* Tx total person years lived beyond age x
* ex remaining life expectancy at age x
* mx per-capita rate of sexual reproduction during the interval [x, x+1)
* lxmx expected number of sexual offspring per original cohort member produced during the interval [x, x+1)

```{r LifeTable, results = 'asis', echo = FALSE}
lifetable <- data.frame(mpm_to_table(
matU,
matF = matR,
matC = NULL,
start = 1L,
xmax = 1000,
lx_crit = 0.01,
radix = 1
))

knitr::kable(lifetable, format = "latex", booktabs = TRUE)

```

\newpage

## Perturbation Matrices

Perturbs elements within a matrix population model and measures the response (sensitivity or elasticity) of the per-capita population growth rate at equilibrium (λ), or, with a user-supplied function, any other demographic statistic.

```{r sensitivity, results = 'asis', echo = FALSE}

senstable <- perturb_matrix(
A,
#pert = 1e-06,
type = "sensitivity",
demog_stat = "lambda",
)

knitr::kable(senstable, format = "latex", booktabs = TRUE, caption =  "Sensitivity Perturbations")

```

```{r elacticity, results = 'asis', echo = FALSE}
elastable <- perturb_matrix(
A,
#pert = 1e-06,
type = "elasticity",
demog_stat = "lambda",
)

knitr::kable(elastable, format = "latex", booktabs = TRUE, caption =  "Elasticity Perturbations")

```

\newpage

## Perturbation Analysis of transition types

Calculates the summed sensitivities or elasticities for various transition types within a matrix population model (MPM), including stasis, retrogression, progression, fecundity, and clonality.

Sensitivities or elasticities are calculated by perturbing elements of the MPM and measuring the response of the per-capita population growth rate at equilibrium (λ), or, with a user-supplied function, any other demographic statistic.

* stasis The sensitivity or elasticity of lambda to stasis.
* retrogression The sensitivity or elasticity of lambda to retrogression.
* progression The sensitivity or elasticity of lambda to progression.
* fecundity The sensitivity or elasticity of lambda to sexual fecundity.
* clonality The sensitivity or elasticity of lambda to clonality.


```{r perturbationtr, results = 'asis', echo = FALSE}

pert_sens <- perturb_trans(
matU,
matR,
matC = NULL,
posU = matU > 0,
posF = matR > 0,
#posC = matC > 0,
exclude_row = NULL,
exclude_col = NULL,
pert = 1e-06,
type = "sensitivity",
demog_stat = "lambda",
)

pert_elas <- perturb_trans(
matU,
matR,
matC = NULL,
posU = matU > 0,
posF = matR > 0,
#posC = matC > 0,
exclude_row = NULL,
exclude_col = NULL,
pert = 1e-06,
type = "elasticity",
demog_stat = "lambda",
)

perturbationtr <- data.frame(matrix(nrow = 5, ncol = 3))
colnames(perturbationtr) <- c("Measure", "Sensitivity", "Elasticity")

perturbationtr$Measure <- c("Stasis", "Retrogression", "Progression", "Fecundity", "Clonality")
perturbationtr$Sensitivity <- c(pert_sens$stasis, pert_sens$retro, pert_sens$progr, pert_sens$fecundity, pert_sens$clonality)
perturbationtr$Elasticity <- c(pert_elas$stasis, pert_elas$retro, pert_elas$progr, pert_elas$fecundity, pert_elas$clonality)

knitr::kable(perturbationtr, format = "latex", booktabs = TRUE)

```

## Perturbation analysis of vital rates

Perturbs lower-level vital rates within a matrix population model and measures the response (sensitivity or elasticity) of the per-capita population growth rate at equilibrium (λ), or, with a usersupplied function, any other demographic statistic. These decompositions assume that all transition rates are products of a stage-specific survival term (column sums of matU) and a lower level vital rate that is conditional on survival (growth, shrinkage, stasis, dormancy, or reproduction). Reproductive vital rates that are not conditional on survival (i.e., within a stage class from which there is no survival) are also allowed.

* survival sensitivity or elasticity of demog_stat to survival
* growth sensitivity or elasticity of demog_stat to growth
* shrinkage sensitivity or elasticity of demog_stat to shrinkage
* fecundity sensitivity or elasticity of demog_stat to sexual fecundity
* clonality sensitivity or elasticity of demog_stat to clonality

```{r perturbationv, results = 'asis', echo = FALSE}

pertv_sens <- perturb_vr(
matU,
matR,
matC = NULL,
pert = 1e-06,
type = "sensitivity",
demog_stat = "lambda"
)


pertv_elas <- perturb_vr(
matU,
matR,
matC = NULL,
pert = 1e-06,
type = "elasticity",
demog_stat = "lambda"
)


perturbationvr <- data.frame(matrix(nrow = 5, ncol = 3))
colnames(perturbationvr) <- c("Measure", "Sensitivity", "Elasticity")

perturbationvr$Measure <- c("Survival", "Growth", "Shrinkage", "Fecundity", "Clonality")
perturbationvr$Sensitivity <- c(pertv_sens$survival, pertv_sens$growth, pertv_sens$shrinkage, pertv_sens$fecundity, pertv_sens$clonality)
perturbationvr$Elasticity <- c(pertv_elas$survival, pertv_elas$growth, pertv_elas$shrinkage, pertv_elas$fecundity, pertv_elas$clonality)

knitr::kable(perturbationvr, format = "latex", booktabs = TRUE)

```

\newpage

## Plot life cycle

```{r lifecycle, fig.cap = "look R can do this for you"}

plot_life_cycle(
A,
title = NULL,
shape = "egg",
fontsize = 10,
nodefontsize = 12,
edgecol = "grey"
)

```

## Age of reproductive maturity

```{r AgeRepro, results = 'asis', echo = FALSE}
prob <- mature_prob(matU, matR, start = 1L)
age <- mature_age(matU, matR, start = 1L)
#mature_distrib(matU, start = 1L, c(FALSE, FALSE, FALSE, TRUE))

AgeReproTable <- data.frame(matrix(, ncol = 2, nrow = 2) )

AgeReproTable[1] <-  c("Prob of reaching mat", "Age at Mat")
AgeReproTable[2] <- c(prob, age)

knitr::kable(AgeReproTable, format = "latex", booktabs = TRUE)

```

## Vital Rates

### Corresponding to separate demographic processes

Derive mean vital rates corresponding to separate demographic processes from a matrix population model. Specifically, this function decomposes vital rates of survival, progression, retrogression, sexual reproduction and clonal reproduction, with various options for weighting and grouping stages of the life cycle.

```{r vitalRates, results = 'asis', echo = FALSE}
vitalrates <- data.frame(vital_rates(
matU,
matR,
matC = NULL,
weights = NULL,
splitStages = "all",
matrixStages = NULL
))

knitr::kable(vitalrates, format = "latex", booktabs = TRUE)
```

### Mean vital rates

```{r meanVitals, results = 'asis', echo = FALSE}
vrsurv <- vr_survival(matU, posU = matU > 0, exclude_col = NULL, weights_col = NULL)
vrgrow <- vr_growth(
matU,
posU = matU > 0,
exclude = NULL,
exclude_row = NULL,
exclude_col = NULL,
weights_col = NULL,
surv_only_na = TRUE
)


vrshrink <- vr_shrinkage(
matU,
posU = matU > 0,
exclude = NULL,
exclude_row = NULL,
exclude_col = NULL,
weights_col = NULL,
surv_only_na = TRUE
)

vrstas <- vr_stasis(
matU,
posU = matU > 0,
exclude = NULL,
weights_col = NULL,
surv_only_na = TRUE
)

# #Not using these
# vr_dorm_enter(matU, posU = matU > 0, dorm_stages, weights_col = NULL)
# vr_dorm_exit(matU, posU = matU > 0, dorm_stages, weights_col = NULL)

VRtable <- data.frame(matrix(, ncol = 2, nrow = 4) )

VRtable[1] <-  c("Rate of Survival", "Rate of Growth", "Rate of shrinkage", "Rate of stasis")
VRtable[2] <- c(vrsurv, vrgrow, vrshrink, vrstas)

knitr::kable(VRtable, format = "latex", booktabs = TRUE)

```

## Survival independent rates

```{r SurvIndRates, results = 'asis', echo = FALSE}

utable <- vr_mat_U(matU, posU = matU > 0, surv_only_na = TRUE)

knitr::kable(utable, format = "latex", booktabs = TRUE, caption = "Survival independent rates - Survival")
```

```{r SurvIndRatesFec, results = 'asis', echo = FALSE}
rtable <- vr_mat_R(matU, matR, posR = matR > 0)
knitr::kable(rtable, format = "latex", booktabs = TRUE, caption = "Survival independent rates - Reproductive")
```

## Stage specific vital rates

```{r vecstage, results = 'asis', echo = FALSE}
vrsurv <- vr_vec_survival(matU, posU = matU > 0, exclude_col = NULL)

vrgrow <- vr_vec_growth(
matU,
posU = matU > 0,
exclude = NULL,
exclude_row = NULL,
exclude_col = NULL,
surv_only_na = TRUE
)

# vr_vec_shrinkage(
# matU,
# posU = matU > 0,
# exclude = NULL,
# exclude_row = NULL,
# exclude_col = NULL,
# surv_only_na = TRUE
# )

vrstas <- vr_vec_stasis(matU, posU = matU > 0, exclude = NULL, surv_only_na = TRUE)

# vr_vec_dorm_enter(matU, posU = matU > 0, dorm_stages)
# vr_vec_dorm_exit(matU, posU = matU > 0, dorm_stages)

vrrepro <- vr_vec_reproduction(
matU,
matR,
posR = matR > 0,
exclude_col = NULL,
weights_row = NULL
)

VRtable2 <- data.frame(matrix(, ncol = 5, nrow = 4) )
colnames(VRtable2) <- c("Stage", "Survival", "Growth", "Stasis", "Reproduction")

VRtable2$Stage <- c(seq(1,4))
VRtable2$Survival <- c(vrsurv)
VRtable2$Growth <- c(vrgrow)
VRtable2$Stasis <- c(vrstas)
VRtable2$Reproduction <- c(vrrepro)

knitr::kable(VRtable2, format = "latex", booktabs = TRUE)
```

\newpage

## Misc
Demetrius’ entropy, Keyfitz’s entropy, generation time, net reproduction, Time to stable stage

Note: method for net reproductive value (R0) from a matrix population model. The net reproduction value (R0) is the mean number of recruits produced during the mean life expectancy of an individual. Note: Can also specify "start" for method but if offspring only arise in stage start, the two methods give the same result.

Note: method for generation time specified as R0,  the time required for a population to increase by a factor of R0 (the net reproductive rate; Caswell (2001), section 5.3.5). Other options: the average parent-offspring age difference (Bienvenu & Legendre (2015)), or the expected age at reproduction for a cohort (Coale (1972), p. 18-19).

``` {r entropy, results = 'asis', echo = FALSE}
ent_d <- entropy_d(lx, mx)
ent_k <- entropy_k(lx, trapeze = FALSE)
generation_time <- gen_time(matU, matR)
nrr <- net_repro_rate(matU, matR, start = 1, method = "generation")
qsd <- qsd_converge(A, start = 1L, conv = 0.01, N = 100000L)

entropytable <- data.frame(matrix(, ncol = 2, nrow = 5) )

entropytable[1] <-  c("Demetrius’ Entropy", "Keyfitz’s Entropy", "Generation Time", "Net Reproduction", "Time to Stable Stage Dist")
entropytable[2] <- c(ent_d, ent_k, generation_time, nrr, qsd)

knitr::kable(entropytable, format = "latex", booktabs = TRUE)

```

