---
title: "CHI Appendix"
author: "Sophie Wulfing, Easton White, Ahilya Sudarshan Kadba"
date: "2023-02-21"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
  bookdown::html_document2:
    fig_caption: yes
    toc: no
  pdf_document:
    toc: no
    fig_caption: yes
bibliography:
- Ch1_Citations.bib
- R_Citations.bib
header-includes: \usepackage{setspace}\doublespacing \usepackage{lineno} \usepackage{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev="cairo_pdf")
#knitr::write_bib look into
setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Mtx_Model")

library(tinytex)
library(matlib)
library(ggplot2)
library(quadprog) #cite
library(gridtext)
library(popbio) #cite
library(knitr)
library(gridExtra)
library(RColorBrewer) 
#library(rPref) #Double check what this is doing. Maybe cite?
library(tidyr)
library(dplyr)
library(ggpubr)
library(bookdown)
library(kableExtra)
library(grid)
library(Rage)


# #Raw data n supplemental. ASK
# in-stage bar graph goes in paper
# in stage management scenarios will definitely be supplemental
# Check out ecological modelling journal
# Write about stable stage cause I don't think you do
# Citations! Also check ahilya table
# APPENDIX!
# MAKE FIG CAPS FOR EVERYTHING AND EDIT FIGS.



observedData <- as.matrix(data.frame(read.csv("RaberinaryBenbow2012Calcs.csv")))
observedStageMatrix <- observedData[1:13, 2:5]
```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

```{r FXNoverallsurvive}
survivability <- function(increase){
    #Calculate survivability for each staege and increase by x%
    NewMtx <- (A * increase) + A
    
    for(i in 1:ncol(NewMtx)-1){
      if(sum(NewMtx[,i]) > 1 | NewMtx[4,4] > 1){
        return(NA)
      }}
    return(NewMtx)
}


#This is stupid but it just forces R to stop writing things as complex numbers
f <- function(x) {
   if (all(Im(z <- zapsmall(x))==0)) as.numeric(z) else x
}

real_check <- function(mtx){
  if(sum(mtx[,1]) > 1 | sum(mtx[,2]) > 1 | sum(mtx[,3]) > 1 | mtx[4,4] > 1){
    return(NA)
  }
  else{return(mtx)}
}

```

```{r QuadProg, echo = FALSE}
#Make Z
z <- matrix(t(observedStageMatrix[2:11,]), ncol = 1, nrow = 40, byrow = TRUE)
#z

#Make M
m <- matrix(, nrow = nrow(z) , ncol = 8)
for(i in 1:(nrow(observedStageMatrix)-3)){
  m[(4 * (i-1) + 1):(4 * i),] <- matrix(c(observedStageMatrix[i,1], 0, 0, 0, 0, 0, observedStageMatrix[i,4], 0,
                                          0, observedStageMatrix[i,1], observedStageMatrix[i,2], 0, 0, 0, 0, 0,
                                          0, 0, 0, observedStageMatrix[i,2], observedStageMatrix[i,3], 0, 0, 0,
                                          0, 0, 0, 0, 0, observedStageMatrix[i,3], 0, observedStageMatrix[i,4]),
                                        nrow = 4, byrow = TRUE)
  #m
}

#Make C
c <- matrix(c(diag(8) * (-1), 
              1, 1, 0, 0, 0, 0, 0, 0,
              0, 0, 1, 1, 0, 0, 0, 0,
              0, 0, 0, 0, 1, 1, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 1), nrow = 12, ncol = 8, byrow = TRUE)
#c
#Make b
b <- matrix(c(rep(0,8), rep(1,4)), nrow = 12, ncol = 1, byrow = TRUE)

#Make P
#p <- matrix(, nrow = 8, ncol = 1, byrow = TRUE) #vector(p1, g1, p2, g2, p3, g3, f4, p4)

#Make mtx G and vector x
g <- t(m) %*% m
f_t <- -(t(-z) %*% m)

qp <- solve.QP(Dmat = g, dvec = f_t, Amat = -t(c), bvec = -t(b))#, factorized = FALSE)


A <- matrix(c(qp$solution[1], 0, 0, qp$solution[7],
                 qp$solution[2], qp$solution[3], 0, 0,
                 0, qp$solution[4], qp$solution[5], 0,
                 0, 0, qp$solution[6], qp$solution[8]), byrow = TRUE, nrow = 4, ncol = 4)
#A
```

# Data

Table \ref{RabData} shows data used to parameterize matrix model from @raberinaryReproductiveCycleOctopus2012. Data was extracted from Figure 7 of this paper using WebPlotDigitizer (https://automeris.io/WebPlotDigitizer/) 

(ref:rabdata) Data collected in Raberinary and Benbow 2012.

```{r RabData, results = 'asis', echo = FALSE, fig.cap = '(ref:rabdata) \\label{RabData}'}

RaberinaryData <- read.csv("RaberinaryBenbow2012Calcs.csv")
#LifeHistory
knitr::kable(RaberinaryData, format = "latex", booktabs = TRUE, caption = "(ref:rabdata) \\label{RabData}") %>%
  kable_styling(latex_options="scale_down")

```


\newpage


# Per-stage management scenarios

## METHODS
Examined how increasing the chance of survival of individuals in each stage class would contribute to population health. This was achieved by isolating the growth ($G_i$) and in-stage survival ($P_i$) for each stage i. We then increased these parameters by 1% and recalculated the overall eigenvalue of the matrix. We then incorporated into different scenarios with different frequencies of fishing restrictions to examine how temporary closures on blue octopus in certain stages would affect the population.

## RESULTS 
Our within stage analysis showed that Stage 1 needed the smallest percent increase in survival to result in overall population growth (Figure \ref{stagemo}). Stage 4 and larval survivability would be the highest needed increase, with stage 4 needing a 25% increase and larval needing a 15% increase in order for the overall population to be stable. Further, when examining the different frequencies of fishing closures, we found that, for any scenario, no closure would be effective if it was less frequent than every other month. As exemplified by the previous analysis, closures focusing on either stage 4 or larval individuals required the most increase in survivorship and highest frequencies of closures in order to result in population growth.


## DISCUSSION
The results of our per-stage analysis showed that focusing on protecting individuals in stage 1 would be the most effective form of management if size could be determined before capture in this fishery. It is a common trend that individuals that survive long larval stages that have high death rates are the most valuable in terms of contributions to overall population growth. However, this is not a realistic management suggestion, as it is difficult to assess the size of octopus before catch, which is often fatal to the individual. This could suggest, however, that the establishment of aquaculture of *Octopus cyanea* could have benefits to the overall population if octopus are reared until passing this first stage of development. However, further research is needed on cephalopod aquaculture in order to be effective and reduce pollutants in the surrounding waters [@jacquetCaseOctopusFarming2023].

(ref:stagemos) Different fishing scenarios based on increasing survivability of one stage.

```{r stagemo, fig.cap = '(ref:stagemos) \\label{stagemo}'}
# I think this is going to be more of an aquaculture conversation than anything else

#Dowble check survivability > 1 check. I don't think you did that correctly
#do stagemo
#have a survivability check in stagemo

stagemo <- function(stage){
  
  stage_it <- data.frame(matrix(ncol=5,nrow=0, dimnames=list(NULL, c("Stage", "Frequency", "PercentChange", "mu", "lambda"))))
  
  for(i in 1:12){
    for(j in 1:100){#FIND OUT HOW TO CONDENSE THESE IF STATEMENTS
      testMtx <- A
      if(stage == 4){
        testMtx[stage, stage] <- A[stage, stage] + A[stage, stage]*(j/100)
      } #input stage 4
      
      else if(stage == 5){
        testMtx[1, stage-1] <- A[1, stage-1] + A[1, stage-1]*(j/100)
      } #input stage 5
      
      else{
      testMtx[,stage] <- A[,stage] + A[,stage]*(j/100)
      } #input other stages
      
      if (!is.na(real_check(testMtx))){
        mu <- ((i-1) * log(eigen(A)$values[1])) + log(eigen(testMtx)$values[1]) / i
        lambda <- exp(mu)
        #print(lambda)
      
        stage_it[nrow(stage_it) + 1,] <- c(stage, i, j/100, f(mu), f(lambda))
       }
    }
  }
  return(stage_it)
}
doesthiswork <- list()
for(i in 1:5){
p1 <-ggplot(stagemo(i), aes((Frequency - 1), PercentChange, fill = lambda)) + geom_tile(width=1) +
  scale_fill_gradient2(
  low = "red",
  mid = "white",
  high = "green",
  #limits = c(1, max(iteration$lambda)),
  midpoint = 1,
  space = "Lab",
  na.value = "grey50"
  )  +
  ylim(0,1) +
  scale_x_continuous(expand=c(0,0), breaks = c(0, 2, 5, 8, 11)) +
  scale_y_continuous(expand=c(0,0)) +
  # scale_x_continuous(name = "Frequency of closure (months)", expand=c(0,0), breaks = c(1, 3, 6, 9, 12)) +
  # scale_y_continuous(name = "Percent change in survival", expand=c(0,0)) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + #, plot.margin = margin(0.5,0.5,0.5,0.5, "cm")) +
  labs(fill = "Eigenvalue")

doesthiswork[[i]] <- p1
}


t <- ggarrange(plotlist = doesthiswork, labels = c("Stage 1", "Stage 2", "Stage 3", "Stage 4","Larval"), label.y = 1.05, label.x = 0.1,  font.label = list(size = 7), common.legend = TRUE, ncol = 2, nrow = 3, legend = "right") + theme(plot.margin = margin(0.1,0.1,0.1,0.1, "cm"))

annotate_figure(t, left = textGrob("Percent change in survival", rot = 90, vjust = 1, gp = gpar(cex = 1)),
                    bottom = textGrob("Months between closures", gp = gpar(cex = 1)))
#do.call(ggarrange, c(doesthiswork[1:5], labels = c("Stage 1", "Stage 2", "Stage 3", "Stage 4","Larval")))

```

\newpage

# Rage package analysis

```{r ragepackage, echo = FALSE}
library(Rage)
matU <- mpm_split(A)$matU
matU[1,1] <- A[1,1]

matR <- mpm_split(A)$matF
matR[1,1] <- 0

#mpm_standardize is the opposite of mpm_split

```

## Age specific Calculations

<!-- (ref:AgeFromStage) Reproduction rates, Survivorship, Survival Probability, and Mortality Hazard for each month of the blue octopus's life history calculated using the Rage Package. -->

<!-- ```{r agefromstage, results = 'asis', echo = FALSE, fig.cap = '(ref:AgeFromStage) \\label{agefromstage}'} -->

<!-- #make mtx with 1_13 rows and each col is vector below -->
<!-- agefromstage <- data.frame(matrix(, nrow = 13, ncol = 5)) -->
<!-- colnames(agefromstage) <- c("Age", "Reproduction", "Survivorship", "SurvivalProbability", "MortalityHazard") -->

<!-- agefromstage$Age <- c(seq(1, 13)) -->

<!-- #Age Specific reproduction -->
<!-- mx <- c(mpm_to_mx(matU, matR, start = 1L, xmax = 1000, lx_crit = 0.01, tol = 1e-04)) -->
<!-- #Age Specific Survivorship -->
<!-- lx <- c(mpm_to_lx(matU, start = 1L, xmax = 1000, lx_crit = 0.01, tol = 1e-04)) -->
<!-- #Age Specific Survival probability -->
<!-- px <- c(mpm_to_px(matU, start = 1L, xmax = 1000, lx_crit = 0.01, tol = 1e-04)) -->
<!-- #Age Specific Mortality Hazard -->
<!-- hx <- c(mpm_to_hx(matU, start = 1L, xmax = 1000, lx_crit = 0.01, tol = 1e-04)) -->


<!-- agefromstage$Reproduction <- mx -->
<!-- agefromstage$Survivorship <- lx -->
<!-- agefromstage$SurvivalProbability <- px -->
<!-- agefromstage$MortalityHazard <- hx -->

<!-- colnames(agefromstage) <- c("Age (month)", "Reproduction", "Survivorship", "Survival Probability", "Mortality Hazard") -->

<!--   knitr::kable(agefromstage, format = "latex", booktabs = TRUE) #%>% -->
<!--   # column_spec(2:3, width = "4.5cm") %>% -->
<!--   # kable_styling(latex_options="scale_down") -->
<!-- ``` -->

Table \ref{LifeTable} shows age specific life-history traits of *Octopus cyanea* as calculated by the Rage package from our matrix.

(ref:lifetable) Life table of *O. cyanea* as calculated from our Lefkovitch matrix.

<!-- * x: age in months at the start of the age interval [x, x+1) -->
<!-- * lx: survivorship, or the proportion of initial cohort surviving to the start of age interval [x, x+1) -->
<!-- * dx: proportion of original cohort dying in the age interval [x, x+1) -->
<!-- * hx: force of mortality (hazard) during the age interval [x, x+1) -->
<!-- * qx: probability of death during the interval [x, x+1) for those entering the interval -->
<!-- * px: probability of survival for the interval [x, x+1) for those entering the interval -->
<!-- * ex: remaining life expectancy at age x -->
<!-- * mx: per-capita rate of sexual reproduction during the interval [x, x+1) -->
<!-- * lxmx: expected number of sexual offspring per original cohort member produced during the interval [x, x+1) -->

```{r LifeTable, results = 'asis', echo = FALSE, fig.cap = '(ref:lifetable) \\label{LifeTable}'}
lifetable <- data.frame(mpm_to_table(
matU,
matF = matR,
matC = NULL,
start = 1L,
xmax = 1000,
lx_crit = 0.01,
radix = 1
))

colnames(lifetable) <- c("Age (months)", "Survivorship", "Proportion of orginal cohort dying", "Mortality hazard", "Probability of death", "Probability of survival", "Remaining life expectancy", "Per-capita reproduction rate", "Expected number of offspring per original cohort number")

knitr::kable(lifetable, format = "latex", booktabs = TRUE)

```

\newpage
























