---
title: "Lefkovitch Method"
author: "Sophie Wulfing"
date: "2/5/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
setwd("C:/Users/sophi/Documents/UNH_Docs/Proposal/Mtx_Model")

library(matlib)
library(ggplot2)

observedData <- as.matrix(data.frame(read.csv("RaberinaryBenbow2012Calcs.csv")))
observedStageMatrix <- observedData[1:13, 2:5]
```

```{r matrix print,results = 'asis', echo = FALSE}
write_matex <- function(x) {
  begin <- "$$\\begin{bmatrix}"
  end <- "\\end{bmatrix}$$"
  X <-
    apply(x, 1, function(x) {
      paste(
        paste(x, collapse = "&"),
        "\\\\"
      )
    })
  writeLines(c(begin, X, end))
}
```
## Note
Adapted from Levkovitch(1963, 1964 a,b) study on cigarette beetle. From page 142 in Caswell book. This method assumes a stable stage dist

```{r initial things, echo = FALSE}
A_reg <- matrix(, nrow = 4, ncol = 5)
SD <- matrix(, nrow = 4, ncol = 1)
Var <- matrix(, nrow = 4, ncol = 1)

#Creating x using first 10 months (11th will be our observed y)
x <- matrix(observedStageMatrix[1:10,], nrow = 10, ncol = 4, byrow = FALSE)
x <- cbind(rep(1, 10), x) #Creates an extra col of 1s at the beginning of x to calculate all the B0s. Going to omit for now because makes the y calcs weird and IDK how to deal with that
```

# Linear Regression
```{r multiple linear regression with matrices}
for (i in 1:nrow(A_reg)){
  #calc each row of mtx and add to A
  y <- matrix(observedStageMatrix[2:11, i])
  arow <- inv((t(x) %*% x)) %*% t(x)  %*% y
  A_reg[i, ] <- arow
  
  #now make vectors for sd and var
  e <- matrix(y - x %*% arow)
  SD[i] <- (t(e) %*% e) / (nrow(y)-nrow(arow))
  Var[i] <- SD[i] * inv(t(x) %*% x)
  
}

```


Calculated A matrix
```{r show A_reg, results = 'asis', echo = FALSE, fig.cap = "Calculated Stage-Based Matrix"}
write_matex(A_reg)
```

SD of matrix
```{r show SD, results = 'asis', echo = FALSE, fig.cap = "Standard Deviation of mtx"}
write_matex(SD)
```

Variance of matrix
```{r show Var, results = 'asis', echo = FALSE, fig.cap = "Variance of mtx"}
write_matex(Var)
```

# Matrix and Predictions
```{r model and predictions, echo = FALSE}
A <- matrix(A_reg[,2:5], nrow = 4, ncol = 4)
pop <- observedStageMatrix[13,]
N <- list()
 
#12 mo projection
for (i in 1:12){
     N[[1]] <- pop
    N[[i+1]] <- A %*% N[[i]]
 }

modeled_data <- t(as.data.frame(do.call(cbind, N)))
colnames(modeled_data) <- c("Stage_1", "Stage_2", "Stage_3", "Stage_4")
modeled_data <- cbind(modeled_data, month = 0:12)

#write.csv(modeled_data, "testing.csv",row.names = TRUE)
modeled_data <- data.frame(modeled_data)

print(ggplot(modeled_data) + 
  geom_line(aes(y = Stage_1, x = month, color = "Stage 1"), size = 1.5) + 
  geom_line(aes(y = Stage_2, x = month, color= "Stage 2"), size = 1.5) +
  geom_line(aes(y = Stage_3, x = month, color = "Stage 3"), size = 1.5) + 
  geom_line(aes(y = Stage_4, x = month, color= "Stage 4"), size = 1.5) +
  scale_color_manual(name = "", values = c("Stage 1" = "#009E73",
                                                "Stage 2" = "#F0E442",
                                                "Stage 3" = "#0072B2",
                                                "Stage 4" = "#D55E00")) +
  ggtitle(paste0(i, " month projection")) +
  xlab("Month") +
  ylab("No. Individuals"))
    

```
\newpage

# Removing Negatives
```{r model remove negs, echo = FALSE}

#Remove Negs from model
A_0 <- A_reg
for(i in 1:nrow(A_0)){
  
  for( j in 2:ncol(A_0)){
    if(A_0[i,j] < 0) {
      A_0[i,j] <- 0
      
    }
  }
}

#Get new SD and Var

SD_0 <- matrix(, nrow = 4, ncol = 1)
Var_0 <- matrix(, nrow = 4, ncol = 1)

for (i in 1:nrow(A_0)){
  #calc each row of mtx and add to A
  y <- matrix(observedStageMatrix[2:11, i])
  arow <- A_0[i,]
  
  #now make vectors for sd and var
  e <- matrix(y - x %*% arow)
  SD_0[i] <- (t(e) %*% e) / (nrow(y)-nrow(A_0))
  Var_0[i] <- SD_0[i] * inv(t(x) %*% x)
  
}

```

A matrix-negatives removed
```{r show A_0, results = 'asis', echo = FALSE, fig.cap = "Calculated Stage-Based Matrix-Negs removed"}
write_matex(A_0)
```

SD of A matrix-negatives removed
```{r show SD_0, results = 'asis', echo = FALSE, fig.cap = "Standard Deviation of mtx-Negs removed"}
write_matex(SD_0)
```

Var of A matrix-negatives removed
```{r show Var_0, results = 'asis', echo = FALSE, fig.cap = "Variance of mtx-Negs removed"}
write_matex(Var_0)
```

```{r model negs removed, echo = FALSE}
  A_0 <- matrix(A_0[,2:5], nrow = 4, ncol = 4)
  pop <- observedStageMatrix[13,]
  N <- list()
 
#12 mo projection
for (i in 1:36){
     N[[1]] <- pop
    N[[i+1]] <- A_0 %*% N[[i]]
 }

modeled_data <- t(as.data.frame(do.call(cbind, N)))
colnames(modeled_data) <- c("Stage_1", "Stage_2", "Stage_3", "Stage_4")
modeled_data <- cbind(modeled_data, month = 0:12)

#write.csv(modeled_data, "testing.csv",row.names = TRUE)
modeled_data <- data.frame(modeled_data)

print(ggplot(modeled_data) + 
  geom_line(aes(y = Stage_1, x = month, color = "Stage 1"), size = 1.5) + 
  geom_line(aes(y = Stage_2, x = month, color= "Stage 2"), size = 1.5) +
  geom_line(aes(y = Stage_3, x = month, color = "Stage 3"), size = 1.5) + 
  geom_line(aes(y = Stage_4, x = month, color= "Stage 4"), size = 1.5) +
  scale_color_manual(name = "", values = c("Stage 1" = "#009E73",
                                                "Stage 2" = "#F0E442",
                                                "Stage 3" = "#0072B2",
                                                "Stage 4" = "#D55E00")) +
  ggtitle(paste0(i, " month projection-Negatives Removed")) +
  xlab("Month") +
  ylab("No. Individuals"))
    

```
\newpage
Next Steps:

1. This model doesn't include the b0 values by adding an extra column of 1's to beginning of X mtx XDONE
2. Try Caswell and Twombly Method (1989) from 143 of Caswell book
3. Try Wood's Quadratic Programming Method 6.2.2 from pg 144 of Caswell book
4. Try to force certain aij's to be zero
5. Look into measuring fecundity and how to deal with maturity btwn T and T+1 (section 6.7)
6. Other tings: Picking timestep, See ch 7 if our timescale is too big (right now this is months for shiggles). If one stage is shorter than timestep, then pi = 0
7. Once you think you have a good model, go back over ch4 for all the stuff you can actually do with that
