---
title: "BauchModel_7"
author: "Sophie Wulfing"
date: "2022-11-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Soc_Ecol_Model")

library(deSolve)
library(tidyverse)
library(ggplot2)
library(gridExtra)
```


```{r Bauch.Model, fig.cap = "Original Bauch Model"}
Bauch.Model <- function (t, y, params) {
  ## first extract the state variables
  P <- y[1] #Fish Pop var
  X <- y[2] #conservationist var

  #TT <- y[5]
  
  ## now extract the parameters
  R <- params["R"] # Net growth/fecundity
  s <- params["s"] # supply and demand
  h <- params["h"] # Harvesting efficiency
  k <- params["k"] # Social learning rate
  w <- params["w"] # conservation costs
  c <- params["c"] # rarity valuation param
  d <- params["d"] # social norm strength
  
  ## now code the model equations
  dPdt <- R * P * (1-P) - (h * P * (1-X))/(P+s)
  dXdt <- k * X * (1-X) * (d*(2*X-1) + 1/(P+c) - w)
  
  ## combine results into a single vector
  dydt <- c(dPdt, dXdt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch <- c(R = 0.06, s = 0.8, h = 0.075, k = 0.17, w = 0.35, c = 0.6, d = 0.5)

times <- seq(from = 0, to = 100, by = 1)
xstart <- c(P = 0.406, X = .24) # See supplemental data for other starting params

ode(
  func=Bauch.Model,
  y=xstart,
  times=times,
  parms=params_Bauch
) %>%
  as.data.frame() -> out1

out1 %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=2)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

```

<!-- Notes: Still worried about interaction term. Don't think this is correct because mechanism isn't specified and we just added interaction terms.  -->

See derivation in notes but this is the new equations (still unsure about interaction terms):
<!-- ##FIX THIS AND EXPLORE FISHING PARAM-->

$\frac{dP_1}{dt} = r_1P_1(1-P_1)-\frac{h_1*P_1(1-X_1)}{P_1 + s_1} + \rho_1 * \frac{P_2 - P_1}{2}$

$\frac{dP_2}{dt} = r_2P_2(1-P_2)-\frac{h_2*P_2(1-X_2)}{P_2 + s_2} + \rho_2 * \frac{P_1 - P_2}{2}$

$\frac{dX_1}{dt} =  k_1X_1(1-X_1) [\frac{1}{P_1+c_1} - w_1 + d_1(2X_1 - 1)] - prop_1X_1(2X_2 - 1)^2 + prop_1(1-X_1)(2X_2 - 1)^2$

$\frac{dX_2}{dt} = k_2x_2(1-X_2)  [\frac{1}{P_2+c_2} - w_2 + d_2(2X_2 - 1)] - prop_2X_2(2X_1 - 1)^2 + prop_2(1-X_2)(2X_1 - 1)^2$


```{r Bauch.Coupled, fig.cap = "New Model with default paramters"}
Bauch.Coupled <- function (t, y, params) {
  ## first extract the state variables
  P_1 <- y[1] #Fish Pop var
  P_2 <- y[2] #Fish Pop var
  X_1 <- y[3] #conservationist var
  X_2 <- y[4] #conservationist var

  #TT <- y[5]
  
  ## now extract the parameters
  R_1 <- params["R_1"] # Net growth/fecundity
  s_1 <- params["s_1"] # supply and demand
  h_1 <- params["h_1"] # Harvesting efficiency
  k_1 <- params["k_1"] # Social learning rate
  w_1 <- params["w_1"] # conservation costs
  c_1 <- params["c_1"] # rarity valuation param
  d_1 <- params["d_1"] # social norm strength
  
  R_2 <- params["R_2"] # Net growth/fecundity
  s_2 <- params["s_2"] # supply and demand
  h_2 <- params["h_2"] # Harvesting efficiency
  k_2 <- params["k_2"] # Social learning rate
  w_2 <- params["w_2"] # conservation costs
  c_2 <- params["c_2"] # rarity valuation param
  d_2 <- params["d_2"] # social norm strength 
  
    roe_1 <- params["roe_1"] # fish diffusion
  roe_2 <- params["roe_2"] # fish diffusion
  prop_1 <- params["prop_1"] # Influence of pop 2 on pop 1
  prop_2 <- params["prop_2"] # Influence of pop 1 on pop 2
  
  ## now code the model equations
  dP_1dt <- R_1 * P_1 * (1-P_1) - (h_1 * P_1 * (1-X_1))/(P_1+s_1) + roe_1 * (P_2 - P_1)/2
  dP_2dt <- R_2 * P_2 * (1-P_2) - (h_2 * P_2 * (1-X_2))/(P_2+s_2) + roe_2 * (P_1 - P_2)/2
  
  dX_1dt <- k_1 * X_1 * (1-X_1) *(1/(P_1 + c_1) - w_1 + d_1*(2*X_1 - 1)) - prop_1 * X_1 * ((2* X_2 - 1)^2) + prop_1 * (1 - X_1) * ((2* X_2 - 1)^2)
  dX_2dt <- k_2 * X_2 * (1-X_2) *(1/(P_2 + c_2) - w_2 + d_2*(2*X_2 - 1)) - prop_2 * X_2 * ((2* X_1 - 1)^2) + prop_2 * (1 - X_2) * ((2* X_1 - 1)^2)
  
  ## combine results into a single vector
  dydt <- c(dP_1dt, dP_2dt, dX_1dt, dX_2dt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch <- c(R_1 = 0.06, s_1 = 0.8, h_1 = 0.075, k_1 = 0.17, w_1 = 0.35, c_1 = 0.6, d_1 = 0.5, roe_1 = 1, prop_1 = 0.1,
                  R_2 = 0.06, s_2 = 0.8, h_2 = 0.075, k_2 = 0.17, w_2 = 0.35, c_2 = 0.6, d_2 = 0.5, roe_2 = 1, prop_2 = 0.1)

times <- seq(from = 0, to = 100, by = 1)
xstart <- c(P_1 = 0.406, X_1 = .24, P_2 = 0.406, X_2 = .24) # See supplemental data for other starting params

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_Bauch
) %>%
  as.data.frame() -> out_Coupled

out_Coupled %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

```


```{r fxnBauch.ParamR}
# params_Default <- data.frame(R = 0.06, s = 0.8, h = 0.075, k = 0.17, w = 0.35, c = 0.6, d = 0.5, roe = 1)
# params_Bauch <- c(R_1 = 0.06, s_1 = 0.8, h_1 = 0.075, k_1 = 0.17, w_1 = 0.35, c_1 = 0.6, d_1 = 0.5, roe_1 = 1,
#                   R_2 = 0.06, s_2 = 0.8, h_2 = 0.075, k_2 = 0.17, w_2 = 0.35, c_2 = 0.6, d_2 = 0.125, roe_2 = 1)

# min <- 0
# max <- 1
# i <- 1
# param <- 8


#Note: in this function, I am changing BOTH populations' paramters. Should I try just changing one?
param_ranges <- function(param, max, min){
  #Generate list of ranges
  new_param <- seq(from=min, to=max, length.out=5)
  plotList <- list()
  
  #Change params to each quartile for graphs 1-5
  for(i in 1:length(new_param)){
    params_Default <- data.frame(R = 0.06, s = .8, h = 0.075, k = 0.17, w = 0.35, c = 0.6, d = 0.5, roe = 1, prop = .1)
    #ASSIGN OLD PARAMS TO BE NEW SEQUENCE
    params_Default[param] <- new_param[i]
    params_Bauch <- c(R_1 = params_Default$R, s_1 = params_Default$s, h_1 = params_Default$h, k_1 = params_Default$k,
                      w_1 = params_Default$w, c_1 = params_Default$c, d_1 = params_Default$d, roe_1 = params_Default$roe,
                      prop_1 =  params_Default$prop,
                      R_2 = params_Default$R, s_2 = params_Default$s, h_2 = params_Default$h, k_2 = params_Default$k,
                      w_2 = params_Default$w, c_2 =params_Default$c, d_2 = params_Default$d, roe_2 = params_Default$roe,
                      prop_2 = params_Default$prop)
    
    times <- seq(from = 0, to = 100, by = 1)
    xstart <- c(P_1 = 0.406, X_1 = .24, P_2 = 0.406, X_2 = .24) #See supplemental for other starting params

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Bauch
    ) %>%
      as.data.frame() -> out_Diff

    p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      labs(x='time (yr)',y='pop')
    plotList[[i]] <- p
  }
  
  #Now assigning the final graph to give the two populations the max and min of the paramters
    params_Default <- data.frame(R = 0.06, s = .8, h = 0.075, k = 0.17, w = 0.35, c = 0.6, d = 0.5, roe = 1, prop = .1)
    #ASSIGN OLD PARAMS TO BE NEW SEQUENCE
    params_Bauch <- c(R_1 = params_Default$R, s_1 = params_Default$s, h_1 = params_Default$h, k_1 = params_Default$k,
                      w_1 = params_Default$w, c_1 = params_Default$c, d_1 = params_Default$d, roe_1 = params_Default$roe,
                      prop_1 =  params_Default$prop,
                      R_2 = params_Default$R, s_2 = params_Default$s, h_2 = params_Default$h, k_2 = params_Default$k,
                      w_2 = params_Default$w, c_2 =params_Default$c, d_2 = params_Default$d, roe_2 = params_Default$roe,
                      prop_2 = params_Default$prop)
    params_Bauch[param] <- max
    params_Bauch[(param + 9)] <- min
    
    times <- seq(from = 0, to = 100, by = 1)
    xstart <- c(P_1 = 0.406, X_1 = .24, P_2 = 0.406, X_2 = .24) #See supplemental for other starting params

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Bauch
    ) %>%
      as.data.frame() -> out_Diff

    p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      labs(x='time (yr)',y='pop')
    plotList[[6]] <- p
  
  
  
  do.call("grid.arrange", c(plotList, ncol=2))
}
  

```

```{r R, fig.cap= "R - Net growth/fecundity, range 0 to 1"}
param_ranges(1, 1, 0)
```

\newpage

```{r s, fig.cap = "S - supply and demand, range 0.1 to 1"}
param_ranges(2, 1, 0.1)
```

\newpage

```{r h, fig.cap = "h - Harvesting efficiency, range 0 to 0.1. Note, default is .075"}
param_ranges(3, 0.1, 0)
```

\newpage

```{r k, fig.cap = "K - Social learning rate 0 to 1"}
param_ranges(4, 1, 0)
```

\newpage

```{r w, fig.cap = "w - conservation costs"}
param_ranges(5, 1, 0)
```

\newpage

```{r c, fig.cap = "c - rarity valuation param"}
param_ranges(6, 1, 0)
```

\newpage

```{r d, fig.cap = "d - social norm strength "}
param_ranges(7, 1, 0)
```

\newpage

```{r roh, fig.cap = "roh - fish diffusion"}
param_ranges(8, 1, 0)
```

\newpage

```{r prop, fig.cap = "prop - Population influence on the other"}
param_ranges(9, 1, 0)
```

```{r fxnparamplane_diffparams}

paramplane <- function(paramx, paramy, maxx, maxy){
  #Create empty dataframe to fill
  paramlist <- c("R", "S", "H", "K", "W", "C", "d", "rho", "prop")
  outcomes <- data.frame(matrix(ncol = 4, nrow = 0))
  colnames(outcomes) <- c("XParameter", "YParameter", "Fish1", "Fish2")
  
  #Dumb but it throws an error if S = 0
  startx <- 0
  starty <- 0
  if(paramx == 2){startx <- .001}
  if(paramy == 2){starty <- .001}
  
  
  #Can comment this part out if using other H vals
  if(paramx == 3){maxx = 0.1}
  if(paramy == 3){maxy = 0.1}
  
  #Create list of all combinations of the two paramters
  paramcombos <- expand.grid(x_parameter = seq(from = startx, to = maxx, length.out = 11),
                        y_parameter = seq(from = starty, to = maxy, length.out = 11))
  
  for(i in 1:nrow(paramcombos)){
  params_Default <- data.frame(R = 0.06, s = .8, h = 0.075, k = 0.17, w = 0.35, c = 0.6, d = 0.5, roe = 1, prop = .1)
  #Change params to be the 
  params_Default[paramx] <- paramcombos$x_parameter[i]
  params_Default[paramy] <- paramcombos$y_parameter[i]
  
  params_Bauch <- c(R_1 = params_Default$R, s_1 = params_Default$s, h_1 = params_Default$h, k_1 = params_Default$k,
                    w_1 = params_Default$w, c_1 = params_Default$c, d_1 = params_Default$d, roe_1 = params_Default$roe,
                    prop_1 =  params_Default$prop,
                    R_2 = params_Default$R, s_2 = params_Default$s, h_2 = params_Default$h, k_2 = params_Default$k,
                    w_2 = params_Default$w, c_2 =params_Default$c, d_2 = params_Default$d, roe_2 = params_Default$roe,
                    prop_2 = params_Default$prop)
    
  times <- seq(from = 0, to = 100, by = 1)
    xstart <- c(P_1 = 0.406, X_1 = .24, P_2 = 0.406, X_2 = .24) #See supplemental for other starting params
    
    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Bauch
    ) %>%
      as.data.frame() -> out_Diff
    
    finalstatus <- out_Diff %>%
      filter(times == 100)
    
    outcomes[i,] <- c(paramcombos$x_parameter[i], paramcombos$y_parameter[i], finalstatus$P_1, finalstatus$P_2)
  }
  
  finalDynamics <- outcomes %>%
    mutate(Result = case_when(
    Fish1 >= .25 & Fish2 >= .25 ~ "Both",
    Fish1 >= .25 & Fish2 < .25 ~ "Only Pop1",
    Fish1 < .25 & Fish2 >= .25 ~ "Only Pop2",
    Fish1 < .25 & Fish2 < .25 ~ "Neither"))
  
cols <- c("green", "orange", "yellow", "black")

#Binned Colors
ggplot(finalDynamics, aes(XParameter, YParameter, fill = Result)) + geom_tile(width=1) +
  scale_fill_manual(
  values = c("Both" = "green",
             "Only Pop1" = "orange",
             "Only Pop2" = "yellow",
             "Neither" = "black")) +
  xlab(paramlist[paramx]) +
  ylab(paramlist[paramy])
}


#paramplane(1, 3, 1, 1)

```

```{r paramplane_testing}

paramplane_testing <- function(boop){

  pl <- list()
  param_tests <- expand.grid(firstParam = seq(from = 1, to = 9, length.out = 9),
                        secondParam = seq(from = 1, to = 9, length.out = 9))

  param_tests <- param_tests %>%
    filter(firstParam != secondParam) %>%
    filter(firstParam == boop)

  for(i in 1:nrow(param_tests)){
    pl[[i]] <- paramplane(param_tests$firstParam[i], param_tests$secondParam[i], 1, 1)
  }

  do.call("grid.arrange", c(pl,  ncol=2))

}

#grid.expand stuff

```

```{r paramplane_testingr, fig.cap = "R parameter planes"}

paramplane_testing(1)

```

\newpage

```{r paramplane_testings, fig.cap = "S parameter planes"}

paramplane_testing(2)

```

\newpage

```{r paramplane_testingh, fig.cap = "h parameter planes"}

paramplane_testing(3)

```

\newpage

```{r paramplane_testingk, fig.cap = "K parameter planes ranging from 0-1"}

paramplane_testing(4)

```


\newpage

```{r paramplane_testingw, fig.cap = "w parameter planes"}

paramplane_testing(5)

```

\newpage

```{r paramplane_testingc, fig.cap = "c parameter planes"}

paramplane_testing(6)

```

\newpage

```{r paramplane_testingd, fig.cap = "d parameter planes"}

paramplane_testing(7)

```

\newpage

```{r paramplane_testingroh, fig.cap = "roh parameter planes"}

paramplane_testing(8)

```

\newpage

```{r paramplane_testingprop, fig.cap = "prop parameter planes"}

paramplane_testing(9)

```

\newpage

```{r fxnparamplane_diffpops}

popplane <- function(param, max){
  #Create empty dataframe to fill
  paramlist <- c("R", "S", "H", "K", "W", "C", "d", "rho", "prop")
  outcomes <- data.frame(matrix(ncol = 4, nrow = 0))
  colnames(outcomes) <- c("Pop1Parameter", "Pop2Parameter", "Fish1", "Fish2")
  
  #Dumb but it throws an error if S = 0
  start <- 0
  if(param == 2){start <- .001}
  
  #Create list of all combinations of the two paramters
  paramcombos <- expand.grid(pop1_parameter = seq(from = start, to = max, length.out = 10),
                        pop2_parameter = seq(from = start, to = max, length.out = 10))
  
  for(i in 1:nrow(paramcombos)){
  params_Default <- data.frame(R = 0.06, s = .8, h = 0.075, k = 0.17, w = 0.35, c = 0.6, d = 0.5, roe = 1, prop = .1)
  #Change params to be the
  
  params_Bauch <- c(R_1 = params_Default$R, s_1 = params_Default$s, h_1 = params_Default$h, k_1 = params_Default$k,
                    w_1 = params_Default$w, c_1 = params_Default$c, d_1 = params_Default$d, roe_1 = params_Default$roe,
                    prop_1 = params_Default$prop,
                    R_2 = params_Default$R, s_2 = params_Default$s, h_2 = params_Default$h, k_2 = params_Default$k,
                    w_2 = params_Default$w, c_2 =params_Default$c, d_2 = params_Default$d, roe_2 = params_Default$roe,
                    prop_2 = params_Default$prop)
  params_Bauch[param] <- paramcombos$pop1_parameter[i]
  params_Bauch[(param + 9)] <- paramcombos$pop2_parameter[i]
    
  times <- seq(from = 0, to = 100, by = 1)
    xstart <- c(P_1 = 0.406, X_1 = .24, P_2 = 0.406, X_2 = .24) #See supplemental for other starting params
    
    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Bauch
    ) %>%
      as.data.frame() -> out_Diff
    
    finalstatus <- out_Diff %>%
      filter(times == 100)
    
    outcomes[i,] <- c(paramcombos$pop1_parameter[i], paramcombos$pop2_parameter[i], finalstatus$P_1, finalstatus$P_2)
  }
  
  finalDynamics <- outcomes %>%
    mutate(Result = case_when(
    Fish1 >= .25 & Fish2 >= .25 ~ "Both",
    Fish1 >= .25 & Fish2 < .25 ~ "Only Pop1",
    Fish1 < .25 & Fish2 >= .25 ~ "Only Pop2",
    Fish1 < .25 & Fish2 < .25 ~ "Neither"))
  
cols <- c("green", "orange", "yellow", "black")

#Binned Colors
ggplot(finalDynamics, aes(Pop1Parameter, Pop2Parameter, fill = Result)) + geom_tile(width=1) +
  scale_fill_manual(
  values = c("Both" = "green",
             "Only Pop1" = "orange",
             "Only Pop2" = "yellow",
             "Neither" = "black")) +
  ggtitle(paste0(paramlist[param], " Parameter plane")) +
  xlab(paste0(paramlist[param], " Population 1")) +
  ylab(paste0(paramlist[param], " Population 2"))
}
```

```{r popplane_testingr, fig.cap = "r population planes"}
popplane(1, 1)

```

```{r popplane_testings, fig.cap = "s population planes"}
popplane(2, 1)

```

```{r popplane_testingh, fig.cap = "h population planes"}
popplane(3, .1)

```

```{r popplane_testingk, fig.cap = "k population planes 0 to 1"}
popplane(4, 1)

```

```{r popplane_testingw, fig.cap = "w population planes"}
popplane(5, 1)

```

```{r popplane_testingc, fig.cap = "c population planes"}
popplane(6, 1)

```

```{r popplane_testingd, fig.cap = "d population planes"}
popplane(7, 1)

```

```{r popplane_testingroh, fig.cap = "roh population planes"}
popplane(8, 1)

```

```{r popplane_testingprop, fig.cap = "prop population planes"}
popplane(9, 1)

```