---
title: "Wulfing_HW08"
author: "Sophie Wulfing"
date: '2022-04-04'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(R2jags)
library(tidyr)
library(dplyr)

setwd("~/UNH_Docs/NR913")

cephs <- read.csv("C:/Users/sophi/Documents/UNH_Docs/Proposal/CNH2_data/cleanData.csv")
```

## Assignment

Fit at least two JAGS models. Fit one model that is “bad”, i.e., has an unacceptable Bayesian p-value (below 0.05 or above  0.95). Then try to fit a different model that is “good”, i.e., has an acceptable Bayesian p-value. I encourage you to fit relatively simple models if using your own data.

### Turn in:

a) an algebraic description of the two models from above. Define all parts of the equation
b) theassociated Bayesian p-value for each

```{r visualize, echo = FALSE}

site_data <- cephs %>%
    group_by(REGION) %>%
    summarise(Count = n()) %>%
    filter(Count > 150)

ceph_sites <- cephs %>%
  filter(REGION %in% site_data$REGION)

ceph_sites2 <- transform(ceph_sites,id=as.numeric(factor(REGION)))

# barplot(site_data$Count)
# 
# tapply(cephs$NO_CAUGHT,cephs$REGION,FUN = mean, na.rm = TRUE)
# 
# # visualize
boxplot(AVG_WT ~ REGION, 
        ylab = "Catch per unit effort (crabs/wk)",
        xlab = "Site",
        outline = TRUE,
        notch = FALSE,
        staplewex = 0.2,
        boxwex = 0.4,
        border = TRUE,
        data = ceph_sites2)
       # col = ccolors[c(5,4)]) #What the counts look like. poisitve, integers, high means
```

```{r badModel, echo = FALSE}

sink("cephs_anova.txt")
cat(" 
    model {
    
    # PRIORS
    for(i in 1:nsites){       #loop for alphas. can also do alpha[1] alpha[2]
    alpha[i] ~ dnorm(0,0.001)
    }
    tau <- 1/sigma^2
    sigma ~ dunif(0,100) #This gives only 1 variance across all sites
    
    # LIKELIHOOD
    for(i in 1:n){
    mass[i] ~ dnorm(mu[i],tau)
    mu[i] <- alpha[site[i]] #Mean will vary by site. 3 alphas, 1 per site. loop through every row of your dataset.
    } # i
    
    # # derived quantitites
    # diff23 <- alpha[3] - alpha[2]

    # # generate new data from the fitted model
    # for(i in 1:n){
    # y.new[i] ~ dnorm(mu[i],tau)   #Generates new data for each run of mcmc
    #  }
    
    } # end of model  
    ",fill = TRUE)
sink()

win.data <- list(mass   = ceph_sites2$AVG_WT,
                 site   = ceph_sites2$id, 
                 n      = nrow(ceph_sites2),
                 nsites = length(unique(ceph_sites2$id)))

# Initial values
inits <- function()list(alpha = rnorm(7), sigma = rlnorm(1))

# Parameters monitored
params <- c("alpha",
            "sigma",
            "Bayes.P"
            #"y.new"
) 

# MCMC settings
ni <- 1000; nt <- 3; nb <- 100; nc <- 3


out <- jags(win.data, inits, params, "cephs_anova.txt", 
            n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, 
            working.directory = getwd())
print(out, dig = 1) 
```


```{r goodModel, echo=FALSE}

```
