---
title: "Wulfing_HW04"
author: "Sophie Wulfing"
date: "2/28/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(R2jags)
library(tidyr)
library(dplyr)

setwd("~/UNH_Docs/NR913")

hares_data <- read.delim("HW3_hareData.txt")

```

## R Question 1

**Using either the Swiss hare dataset or your own data, run a Bayesian t-test OR a simple linear 
regression using JAGS.**

Hi sorry, I know you said to use our own data but I'm still in the middle of a data pull. I'll hopefully have some for next assignment but for now, hares. Also turns out elevation is not a very good predictor of hare density lol.

```{r q1}
hares <- hares_data %>% drop_na(elevation) %>% drop_na(mean.density)
  

#sink("HW4_model.txt") #Had to comment this out so markdown would knit
 cat("model { #always start JAGS with this model line

     # Priors-all things we don't know
     beta0 ~ dnorm(0,0.01)		# precision inverse of variance. This means huge variance
     beta1 ~ dnorm(0,0.01)
     precision <- 1 / variance	#Priors are unknown. We only know mass and svl
     variance <- sigma^2
     sigma ~ dunif(0,15) #I just kept 15 cause that's what I've been using
     #No prior for mew. We will calc further down. We've covered mew using priors for b0 and b1


     # Likelihood
     for(i in 1:nobs){
     mean.density[i] ~ dnorm(mu[i], precision) #mass - > density

     mu[i] <- beta0 + beta1 * elevation[i] #svl-> elevation

     } # i loop

     } # end of the model.
     ",fill=TRUE)
#sink() #I had to comment this out to get markdown to knit

# Bundle data
win.data <- list(mean.density = hares$mean.density, 
                 elevation = hares$elevation, 
                 nobs = nrow(hares))

# Function to generate starting values aka initial values. Supply init vals
inits <- function()list(beta0 = rnorm(1), 
                        sigma = runif(1, 0, 15))

# Parameters to be monitored (= to estimate)
params <- c("beta0", 
            "beta1", 
            "sigma")

# MCMC settings
nc <- 3
ni <- 1000
nb <- 1
nt <- 1

out <- jags(win.data, inits, params, "HW4_model.txt", n.chains = nc, 
            n.thin = nt, n.iter = ni, n.burnin = nb, working.directory = getwd())

print(out, dig =2) 

m <- lm(mean.density ~ elevation, hares)
summary(m)

elevation_p <- seq(min(hares$elevation), max(hares$elevation),length.out = length(hares$elevation))

elevation_preds <- m$coefficients[1] + m$coefficients[2]*elevation_p



plot(hares$mean.density ~ hares$elevation, 
     ylab = "Hare Density", 
     xlab = "Elevation",
     pch = 20)
lines(elevation_preds ~ elevation_p)
abline(a =out$BUGSoutput$mean$beta0, 
       b = out$BUGSoutput$mean$beta1,
       lwd = 4,lty = 3, col = "red")
# plot. Bayes vs linear mocel from above. Black is truth, blue is lm, red is bayes

```

\newpage

## BONUS

**Fit a means parameterization of a t-test in JAGS, either to the Swiss hares data, 
simulation data, or your own data. Hint: see page 120-121 of Kéry 2010 regarding double indexing syntax**

Ok I have no idea how to actually do this. I figured out the alpha indexing thing but couldn't get the equation quite right until class today. However, when I set up the matrix (d) it still down't look right. Is it a model issue or am I just setting up that matrix incorrectly?

```{r bonus}
bonus_hares <- hares_data %>% 
        drop_na(landuse) %>% 
        drop_na(mean.density )

for(i in 1:nrow(bonus_hares)){
        if(bonus_hares$landuse[i] == "arable"){
                bonus_hares$x[i] <- 1
        }
        else{ # if(bonus_hares$landuse[i] == "grass")
                bonus_hares$x[i] <- 2
        }
}

# bonus_hares
# 
# data_sum <- bonus_hares  %>%
#         group_by(landuse) %>%
#         summarise_at(vars(mean.density), list(name = mean, sd))
# 
# mu1 <- as.numeric(data_sum[1,2]) #Arable land mean
# mu2 <- as.numeric(data_sum[2,2]) #Grassland mean
# 
# sigma1 <- as.numeric(data_sum[1,3]) #Arable land sd
# sigma2 <- as.numeric(data_sum[2,3]) #Grassland sd

#sink("HW4_bonusModel.txt")
cat(" 
    model {

    # PRIORS
    for(j in 1:nsites){
      alpha[j] ~ dnorm(0,0.001)
      tau[j] <- 1/sigma[j]^2
      sigma[j] ~ dunif(0,10)
    }

    # LIKELIHOOD
    for(i in 1:n){
      y[i] ~ dnorm(mu[i],tau[x[i]])

      #mu[i] <- alpha0 + alpha1*x[i]
      mu[i] <-  alpha[x[i]]

    }
    
    diff12 <- alpha[2] - alpha[1]

    for(i in 1:nsites){
    avg[i] <- (alpha[i]*stdev_data) + mean_data
    stan_dev[i] <- (sigma[i]*stdev_data) 

    }
    } # end of model  
    ",fill = TRUE)
#sink()

win.data <- list(y= bonus_hares$mean.density,
                 n = nrow(bonus_hares),
                 x = bonus_hares$x,
                 nsites = length(unique(bonus_hares$x)),#This is just going to be 2
                 stdev_data = sd(bonus_hares$mean.density),
                 mean_data = mean(bonus_hares$mean.density))

# Initial values
inits <- function()list(alpha = rnorm(length(unique(bonus_hares$x))), 
                        sigma = rlnorm(length(unique(bonus_hares$x))))

# Parameters monitored
params <- c("alpha",
            "sigma",
            "diff12",
            "avg",
            "stan_dev"
) 

# MCMC settings
ni <- 10000; nt <- 1; nb <- 1000; nc <- 3


out <- jags(win.data, inits, params, "HW4_bonusModel.txt", 
            n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb, 
            working.directory = getwd())
print(out, dig = 2)

# FIX THISsee the full columns of the design matrix and linear algebra
d <- data.frame(density = round(bonus_hares$mean.density,1),
           int = rep(1,length(bonus_hares$mean.density)),
           alpha1 = round(rep(out$BUGSoutput$mean$alpha[1], length(bonus_hares$mean.density)),1),
           landuse = bonus_hares$x,
           alpha2 = round(rep(out$BUGSoutput$mean$alpha[2], length(bonus_hares$mean.density)),1))
           #resid = round(out$BUGSoutput$mean$resid,1))

print(d)

```

