---
title: "BibExample"
author: "Sophie Wulfing"
date: "11/9/2021"
output: pdf_document
bibliography: example_bib.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ways to use Markdown like Latex:

I was reading this awesome paper [@Shea2014]. Citer package can help you search for citations in zotero. Can export bib file from zotero.

```{r cars}
summary(cars)
plot(cars$speed, cars$dist)

my_mod <- lm(cars$dist~cars$speed)
my_mod_summary <- summary(my_mod)
abline(my_mod)

my_mod_summary$coefficients #str() gives you structure (i.e. you can exrract residuals and coefficients form summaries of linear regression)

my_mod_summary$coefficients[2, 4] #extract info

avg_dist <- mean(cars$dist)
```


I like the number `r 2+2`

I found that the average stopping distance was `r round(avg_dist)`

Got to settings, Chunk output in console to make it stop plotting in line.

Distance was `r ifelse(my_mod_summary$coefficients[2, 4] < 0.05, '', 'not')` significantly (p=value = `r my_mod_summary$coefficients[2, 4]`) correlated with speed.



```{r pressure, echo=FALSE, fig.cap = "blah blah blee blah blah blah \\label{fig:pressure}"}
plot(pressure)
```


As you can see, in Figure \ref{fig:pressure}
\newpage