---
title: "Reviewer1_MovementQ"
author: "Sophie Wulfing"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: true
    toc: false
    lof: true
    lot: true
    keep_tex: true
  pdf_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev="cairo_pdf", cache = TRUE)

library(deSolve)
library(tidyverse)
library(ggplot2)
library(gridExtra)
#library(kableExtra)
library(FME)
library(ggpubr)
library(gridtext)

#setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Soc_Ecol_Model")

params_list <- data.frame(r_1 = 0.16, s_1 = 0.8, h_1 = .25, k_1 = 0.17, w_1 = 1.44, c_1 = 0.5, d_1 = 0.3, m_1 = 0.01, rho_1 = 0.01,
                          r_2 = 0.16, s_2 = 0.8, h_2 = .25, k_2 = 0.17, w_2 = 1.44, c_2 = 0.5, d_2 = 0.3, m_2 = 0.01, rho_2 = 0.01)


cbbPalette <- c("#E69F00", "#009E73", "#567de9", "#D55E00") #F_1, F_2, X_1, X_2

# #WORKED - MINIMUM GROWTH RATE USED IN THAMPI
# params_dispersal <- data.frame(r_1 = 0.16, s_1 = 0.8, h_1 = 0.5, k_1 = 0.17, w_1 = 1.44, c_1 = 0.6, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.16, s_2 = 0.8, h_2 = 0.5, k_2 = 0.17, w_2 = 1.44, c_2 = 0.6, d_2 = 0.3, m_2 = 0, rho_2 = 0)

#Extensions of Bauch Default but with my params I want changed. Goal: get this to oscillate
# params_dispersal <- data.frame(r_1 = 0.35, s_1 = 0.8, h_1 = 0.5, k_1 = 0.17, w_1 = 1, c_1 = 1.68, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 0.17, w_2 = 1, c_2 = 1.68, d_2 = 0.3, m_2 = 0, rho_2 = 0)


#Extension of Bauch Default
# params_dispersal <- data.frame(r_1 = 0.06, s_1 = 0.8, h_1 = 0.1, k_1 = 0.17, w_1 = 1, c_1 = 0.6, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.06, s_2 = 0.8, h_2 = 0.1, k_2 = 0.17, w_2 = 1, c_2 = 0.6, d_2 = 0.3, m_2 = 0, rho_2 = 0)


xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)
timespan <- seq(from = 0, to = 1000, by = 1)

```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

```{r DefaultParamTable, results = 'asis', echo = FALSE}


giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Definition")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "m", "$\\rho$")
giveparams$Population_1 <- c(params_list[1:9])
giveparams$Population_2 <- c(params_list[10:18])
colnames(giveparams) <- c("Parameter", "Population 1", "Population 2", "Definition")

giveparams$Definition <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Rate of sampling opinions or social interaction",
                    "Conservation cost",
                    "Rarity valuation",
                    "Strength of social influence (within population)",
                    "Fish movement (from opposite patch)",
                    "Strength of social influence (from opposite population)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "(ref:defaultparamtable) \\label{DefaultParamTable}")

```

```{r Bauch.Coupled, warning = FALSE}


##Maybe see what proportion of people can actually be fishing in this system to make it sustainable

Bauch.Coupled <- function (t, y, params) {
  ## first extract the state variables
  F_1 <- y[1] #Fish Pop var
  F_2 <- y[2] #Fish Pop var
  X_1 <- y[3] #conservationist var
  X_2 <- y[4] #conservationist var

  #TT <- y[5]
  
  ## now extract the parameters
  r_1 <- params["r_1"] # Net growth/fecundity
  s_1 <- params["s_1"] # supply and demand
  h_1 <- params["h_1"] # Harvesting efficiency
  k_1 <- params["k_1"] # Social learning rate
  w_1 <- params["w_1"] # conservation costs
  c_1 <- params["c_1"] # rarity valuation param
  d_1 <- params["d_1"] # social norm strength
  m_1 <- params["m_1"] # Inflow of fish from pop 2 to pop 1
  rho_1 <- params["rho_1"] # Influence of pop 2 on pop 1
  
  r_2 <- params["r_2"] # Net growth/fecundity
  s_2 <- params["s_2"] # supply and demand
  h_2 <- params["h_2"] # Harvesting efficiency
  k_2 <- params["k_2"] # Social learning rate
  w_2 <- params["w_2"] # conservation costs
  c_2 <- params["c_2"] # rarity valuation param
  d_2 <- params["d_2"] # social norm strength 
  m_2 <- params["m_2"] # Inflow of fish from pop 1 to pop 2
  rho_2 <- params["rho_2"] # Influence of pop 1 on pop 2
  
  ## now code the model equations
  dF_1dt <-  r_1 * F_1 * (1-F_1) - (h_1 * F_1 * (1-X_1))/(F_1+s_1) - m_2 * F_1 + m_1 * F_2
  dF_2dt <-  r_2 * F_2 * (1-F_2) - (h_2 * F_2 * (1-X_2))/(F_2+s_2) - m_1 * F_2 + m_2 * F_1
  
  dX_1dt <- k_1 * X_1 * (1-X_1) *(1/(F_1 + c_1) - w_1 + d_1*(2*X_1 - 1) + rho_1 * (2*X_2 - 1))
  dX_2dt <- k_2 * X_2 * (1-X_2) *(1/(F_2 + c_2) - w_2 + d_2*(2*X_2 - 1) + rho_2 * (2*X_1 - 1))
  
  ## combine results into a single vector
  dydt <- c(dF_1dt, dF_2dt, dX_1dt, dX_2dt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch2 <- params_list

times <- timespan
xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_Bauch2
) %>%
  as.data.frame() -> out_Coupled

out_Coupled %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  scale_colour_manual(values=cbbPalette) +
  labs(x='Time (yr)',y='Population') +
  theme(legend.title=element_blank()) +
  scale_discrete_manual(values=cbbPalette, aesthetics = "colour",name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2])))


```

(ref:movementbothrho) In graphs a), b), and c), both $\rho_1$ and $\rho_2$ were set to 0.01, 0.25, and 0.5, respectively. The corresponding graphs show the dynamics of these models with the new parameterizations. d), e), and f) show the changes in model dynamics when $m_2$ is held at 0.01 and only $m_1$ (the movement of resources from patch 2 to patch 1) is increased by 0.01, 0.05, and 0.1, respectively.


```{r MovementBothRho, warning = FALSE, fig.cap='(ref:movementbothrho) \\label{MovementBothRho}'}

mparams <- c(0.01, 0.05, 0.1)

rhoparams <- c(0.01, 0.25, 0.5)

plotList <- list()
k <- 0
titlesList <- c("a)", "b)", "c)", "d)", "e)", "f)")

for(i in 1:length(rhoparams)){
  k <- k+1
  params_Default <- params_list
  params_Default[9] <- rhoparams[i]
  params_Default[18] <- rhoparams[i]

  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
  if(k == 1) {
    p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = "Both Rho", subtitle = titlesList[k]) +
      theme(plot.title = element_text(hjust = .5))
    plotList[[k]] <- p
    
  }else{
    p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = NULL, subtitle = titlesList[k])
    plotList[[k]] <- p
  }
}

for(i in 1:length(mparams)){
  k <- k+1
  params_Default <- params_list
  params_Default[8] <- mparams[i]
  #params_Default[17] <- mparams[i]

  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff

    
    if(k == 4) {
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        labs(x='Time (yr)',y='Population') +
      ggtitle(label = "Asymmetrical Movement", subtitle = titlesList[k]) +
      theme(plot.title = element_text(hjust = .5))
      plotList[[k]] <- p
    
    }else if(k == 6){
      p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = NULL, subtitle = titlesList[k]) +
      scale_discrete_manual(values=cbbPalette, aesthetics = "colour",name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2])))
    plotList[[k]] <- p
    
    }else{
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        labs(x='Time (yr)',y='Population') +
        ggtitle(label = NULL, subtitle = titlesList[k])
      plotList[[k]] <- p
    }
  
}

#function to extract the legend of a ggplot; source:
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs

get_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}


#arranging the legend and plots in a grid:

p_legend <- get_legend(plotList[[6]])

grid.arrange(arrangeGrob(plotList[[1]] + theme(legend.position="none"),
                         plotList[[4]] + theme(legend.position="none"),
                         plotList[[2]] + theme(legend.position="none"),
                         plotList[[5]] + theme(legend.position="none"),
                         plotList[[3]] + theme(legend.position="none"),
                         plotList[[6]] + theme(legend.position="none"),
                         nrow=3),
             p_legend,
             ncol=2,widths=c(2, 0.3))

```

(ref:movementonerho) In graphs a), b), and c), $\rho_1$ was set to 0.01, 0.5, and 1, respectively. The corresponding graphs show the dynamics of these models with the new parameterizations. d), e), and f) show the changes in model dynamics when $m_2$ is held at 0.01 and only $m_1$ (the movement of resources from patch 2 to patch 1) is increased by 0.01, 0.05, and 0.1, respectively.


```{r MovementOneRho, warning = FALSE, fig.cap='(ref:movementonerho) \\label{MovementOneRho}'}

mparams <- c(0.01, 0.05, .1)

rhoparams <- c(0.01, 0.5, 1)

plotList <- list()
k <- 0
titlesList <- c("a)", "b)", "c)", "d)", "e)", "f)")

for(i in 1:length(rhoparams)){
  k <- k+1
  params_Default <- params_list
  params_Default[9] <- rhoparams[i]
  #params_Default[18] <- rhoparams[i]

  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
  if(k == 1) {
    p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = "Rho1", subtitle = titlesList[k]) +
      theme(plot.title = element_text(hjust = .5))
    plotList[[k]] <- p
    
  }else{
    p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = NULL, subtitle = titlesList[k])
    plotList[[k]] <- p
  }
}

for(i in 1:length(mparams)){
  k <- k+1
  params_Default <- params_list
  params_Default[8] <- mparams[i]
  #params_Default[17] <- mparams[i]

  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
    
    if(k == 4) {
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        labs(x='Time (yr)',y='Population') +
      ggtitle(label = "Asymmetrical Movement", subtitle = titlesList[k]) +
      theme(plot.title = element_text(hjust = .5))
      plotList[[k]] <- p
    
    }else if(k == 6){
      p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = NULL, subtitle = titlesList[k]) +
      scale_discrete_manual(values=cbbPalette, aesthetics = "colour",name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2])))
    plotList[[k]] <- p
    
    }else{
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        labs(x='Time (yr)',y='Population') +
        ggtitle(label = NULL, subtitle = titlesList[k])
      plotList[[k]] <- p
    }
  
}

#function to extract the legend of a ggplot; source:
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs

get_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}


#arranging the legend and plots in a grid:

p_legend <- get_legend(plotList[[6]])

grid.arrange(arrangeGrob(plotList[[1]] + theme(legend.position="none"),
                         plotList[[4]] + theme(legend.position="none"),
                         plotList[[2]] + theme(legend.position="none"),
                         plotList[[5]] + theme(legend.position="none"),
                         plotList[[3]] + theme(legend.position="none"),
                         plotList[[6]] + theme(legend.position="none"),
                         nrow=3),
             p_legend,
             ncol=2,widths=c(2, 0.3))

```

\newpage

# Looking at rho params in symmetrical case

(ref:influenceasym) Listening to yourself ($d_1$) vs other pop ($\rho_1$)

```{r influenceAsym, warning = FALSE, fig.cap = '(ref:influenceasym) \\label{influenceAsym}'}

rhoparams <- c(0.01, 0.25, 0.5)
dparams <- c(0.01, 0.25, 0.5)

plotList <- list()
k <- 0

for(i in 1:length(dparams)){
  params_Default <- params_list
  params_Default[7] <- dparams[i]
  #params_Default[16] <- dparams[i]
  
  for(j in 1:length(rhoparams)){
    k <- k+1
    params_Default[9] <- rhoparams[j]
    #params_Default[18] <- rhoparams[j]
  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
    if(k ==1){
      p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      scale_x_discrete(name = "", limits = c(0, 500, 950)) +
      scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
      # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
      # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
      scale_discrete_manual(values=cbbPalette, aesthetics = "colour", name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2]))) +
      ggtitle(paste("\u03B4 = ", params_Default[7], "\u03C1 = ", params_Default[9])) +
      theme(plot.title = element_text(size=7))
    plotList[[k]] <- p
    
    }else{
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        scale_x_discrete(name = "", limits = c(0, 500, 950)) +
        scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
        # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
        # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
        #theme(legend.position = "none") +
        ggtitle(paste("\u03B4 = ", params_Default[7], "\u03C1 = ", params_Default[9])) +
        theme(plot.title = element_text(size=7))
      plotList[[k]] <- p
    }
  }
}

yleft = richtext_grob("Population", rot=90)

bottom = richtext_grob(
  text = 'Time (year)'
)
  

  p_legend <- get_legend(plotList[[1]])

grid.arrange(arrangeGrob(plotList[[1]] + theme(legend.position="none"),
                         plotList[[2]] + theme(legend.position="none"),
                         plotList[[3]] + theme(legend.position="none"),
                         plotList[[4]] + theme(legend.position="none"),
                         plotList[[5]] + theme(legend.position="none"),
                         plotList[[6]] + theme(legend.position="none"),
                         plotList[[7]] + theme(legend.position="none"),
                         plotList[[8]] + theme(legend.position="none"),
                         plotList[[9]] + theme(legend.position="none"),
                         nrow=3, left = yleft, bottom = bottom),
             p_legend, 
             ncol=2,widths=c(2, 0.3))

```