---
title: "Review1_NewFigs"
author: "Sophie Wulfing"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: true
    toc: false
    lof: true
    lot: true
    keep_tex: true
  pdf_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev="cairo_pdf", cache = TRUE)

library(deSolve)
library(tidyverse)
library(ggplot2)
library(gridExtra)
#library(kableExtra)
library(FME)
library(ggpubr)
library(gridtext)

#setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Soc_Ecol_Model")

params_list <- data.frame(r_1 = 0.16, s_1 = 0.8, h_1 = .25, k_1 = 0.17, w_1 = 1.44, c_1 = 0.5, d_1 = 0.3, m_1 = 0.01, rho_1 = 0.01,
                          r_2 = 0.16, s_2 = 0.8, h_2 = .25, k_2 = 0.17, w_2 = 1.44, c_2 = 0.5, d_2 = 0.3, m_2 = 0.01, rho_2 = 0.01)


cbbPalette <- c("#E69F00", "#009E73", "#567de9", "#D55E00") #F_1, F_2, X_1, X_2

# #WORKED - MINIMUM GROWTH RATE USED IN THAMPI
# params_dispersal <- data.frame(r_1 = 0.16, s_1 = 0.8, h_1 = 0.5, k_1 = 0.17, w_1 = 1.44, c_1 = 0.6, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.16, s_2 = 0.8, h_2 = 0.5, k_2 = 0.17, w_2 = 1.44, c_2 = 0.6, d_2 = 0.3, m_2 = 0, rho_2 = 0)

#Extensions of Bauch Default but with my params I want changed. Goal: get this to oscillate
# params_dispersal <- data.frame(r_1 = 0.35, s_1 = 0.8, h_1 = 0.5, k_1 = 0.17, w_1 = 1, c_1 = 1.68, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 0.17, w_2 = 1, c_2 = 1.68, d_2 = 0.3, m_2 = 0, rho_2 = 0)


#Extension of Bauch Default
# params_dispersal <- data.frame(r_1 = 0.06, s_1 = 0.8, h_1 = 0.1, k_1 = 0.17, w_1 = 1, c_1 = 0.6, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.06, s_2 = 0.8, h_2 = 0.1, k_2 = 0.17, w_2 = 1, c_2 = 0.6, d_2 = 0.3, m_2 = 0, rho_2 = 0)


xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)
timespan <- seq(from = 0, to = 1000, by = 1)

```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

```{r DefaultParamTable, results = 'asis', echo = FALSE}


giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Definition")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "m", "$\\rho$")
giveparams$Population_1 <- c(params_list[1:9])
giveparams$Population_2 <- c(params_list[10:18])
colnames(giveparams) <- c("Parameter", "Population 1", "Population 2", "Definition")

giveparams$Definition <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Rate of sampling opinions or social interaction",
                    "Conservation cost",
                    "Rarity valuation",
                    "Strength of social influence (within population)",
                    "Fish movement (from opposite patch)",
                    "Strength of social influence (from opposite population)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "(ref:defaultparamtable) \\label{DefaultParamTable}")

```

```{r Bauch.Coupled, warning = FALSE}


##Maybe see what proportion of people can actually be fishing in this system to make it sustainable

Bauch.Coupled <- function (t, y, params) {
  ## first extract the state variables
  F_1 <- y[1] #Fish Pop var
  F_2 <- y[2] #Fish Pop var
  X_1 <- y[3] #conservationist var
  X_2 <- y[4] #conservationist var

  #TT <- y[5]
  
  ## now extract the parameters
  r_1 <- params["r_1"] # Net growth/fecundity
  s_1 <- params["s_1"] # supply and demand
  h_1 <- params["h_1"] # Harvesting efficiency
  k_1 <- params["k_1"] # Social learning rate
  w_1 <- params["w_1"] # conservation costs
  c_1 <- params["c_1"] # rarity valuation param
  d_1 <- params["d_1"] # social norm strength
  m_1 <- params["m_1"] # Inflow of fish from pop 2 to pop 1
  rho_1 <- params["rho_1"] # Influence of pop 2 on pop 1
  
  r_2 <- params["r_2"] # Net growth/fecundity
  s_2 <- params["s_2"] # supply and demand
  h_2 <- params["h_2"] # Harvesting efficiency
  k_2 <- params["k_2"] # Social learning rate
  w_2 <- params["w_2"] # conservation costs
  c_2 <- params["c_2"] # rarity valuation param
  d_2 <- params["d_2"] # social norm strength 
  m_2 <- params["m_2"] # Inflow of fish from pop 1 to pop 2
  rho_2 <- params["rho_2"] # Influence of pop 1 on pop 2
  
  ## now code the model equations
  dF_1dt <-  r_1 * F_1 * (1-F_1) - (h_1 * F_1 * (1-X_1))/(F_1+s_1) - m_2 * F_1 + m_1 * F_2
  dF_2dt <-  r_2 * F_2 * (1-F_2) - (h_2 * F_2 * (1-X_2))/(F_2+s_2) - m_1 * F_2 + m_2 * F_1
  
  dX_1dt <- k_1 * X_1 * (1-X_1) *(1/(F_1 + c_1) - w_1 + d_1*(2*X_1 - 1) + rho_1 * (2*X_2 - 1))
  dX_2dt <- k_2 * X_2 * (1-X_2) *(1/(F_2 + c_2) - w_2 + d_2*(2*X_2 - 1) + rho_2 * (2*X_1 - 1))
  
  ## combine results into a single vector
  dydt <- c(dF_1dt, dF_2dt, dX_1dt, dX_2dt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch2 <- params_list

times <- timespan
xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_Bauch2
) %>%
  as.data.frame() -> out_Coupled

# out_Coupled %>%
#   gather(variable,value,-time) %>%
#   ggplot(aes(x=time,y=value,color=variable))+
#   geom_line(size=1)+
#   theme_classic()+
#   scale_colour_manual(values=cbbPalette) +
#   labs(x='Time (yr)',y='Population') +
#   theme(legend.title=element_blank()) +
#   scale_discrete_manual(values=cbbPalette, aesthetics = "colour",name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2])))


```


# NEW FIG

Figure \ref{MovementBothRho} is going to be a new addition to the manuscript. This is basically showing that increases in rho delays decision making but a high increase will eventually lead to collapse. This can also relate to reviewers point about delayed decision making and timescales of conservation decisions

(ref:movementbothrho) In graphs a), b), and c), both $\rho_1$ and $\rho_2$ were set to 0.01, 0.25, and 0.5, respectively. The corresponding graphs show the dynamics of these models with the new parameterizations. d), e), and f) show the changes in model dynamics when $\rho_2$ is held at 0.01 and only $\rho_1$, the social influence of human population 1 onto human population 2 is increased by 0.01, 0.25, and 0.5, respectively.


```{r MovementBothRho, warning = FALSE, fig.cap='(ref:movementbothrho) \\label{MovementBothRho}'}

#mparams <- c(0.01, 0.05, 0.1)

rhoparams <- c(0.01, 0.25, 0.5)

plotList <- list()
k <- 0
titlesList <- c("a)", "b)", "c)", "d)", "e)", "f)")

for(i in 1:length(rhoparams)){
  k <- k+1
  params_Default <- params_list
  params_Default[9] <- rhoparams[i]
  params_Default[18] <- rhoparams[i]

  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
  if(k == 1) {
    p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = "", subtitle = titlesList[k]) +
      theme(plot.title = element_text(hjust = .5))
    plotList[[k]] <- p
    
  }else{
    p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = NULL, subtitle = titlesList[k])
    plotList[[k]] <- p
  }
}

for(i in 1:length(rhoparams)){
  k <- k+1
  params_Default <- params_list
  params_Default[9] <- rhoparams[i]
  #params_Default[18] <- rhoparams[i]

  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff

    
    if(k == 4) {
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        labs(x='Time (yr)',y='Population') +
      ggtitle(label = "", subtitle = titlesList[k]) +
      theme(plot.title = element_text(hjust = .5))
      plotList[[k]] <- p
    
    }else if(k == 6){
      p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      labs(x='Time (yr)',y='Population') +
      ggtitle(label = NULL, subtitle = titlesList[k]) +
      scale_discrete_manual(values=cbbPalette, aesthetics = "colour",name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2])))
    plotList[[k]] <- p
    
    }else{
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        labs(x='Time (yr)',y='Population') +
        ggtitle(label = NULL, subtitle = titlesList[k])
      plotList[[k]] <- p
    }
  
}

#function to extract the legend of a ggplot; source:
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs

get_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}


#arranging the legend and plots in a grid:

p_legend <- get_legend(plotList[[6]])

grid.arrange(arrangeGrob(plotList[[1]] + theme(legend.position="none"),
                         plotList[[4]] + theme(legend.position="none"),
                         plotList[[2]] + theme(legend.position="none"),
                         plotList[[5]] + theme(legend.position="none"),
                         plotList[[3]] + theme(legend.position="none"),
                         plotList[[6]] + theme(legend.position="none"),
                         nrow=3),
             p_legend,
             ncol=2,widths=c(2, 0.3))

```


# REPLACEMENT FIG

Make sure you change params on this document code

You could also possibly put the BothRhos document into appendix to show delayed decision making

```{r NewParamTable, results = 'asis', echo = FALSE}

params_list <- data.frame(r_1 = 0.40, s_1 = 0.8, h_1 = 0.25, k_1 = 1.014, w_1 = 0.2, c_1 = 1.5, d_1 = 0.5, m_1 = 0.2, rho_1 = 0.5,
                               r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 1.014, w_2 = 0.35, c_2 = 1.5, d_2 = 0.5, m_2 = 0.2, rho_2 = 0.1)


giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Definition")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "m", "$\\rho$")
giveparams$Population_1 <- c(params_list[1:9])
giveparams$Population_2 <- c(params_list[10:18])
colnames(giveparams) <- c("Parameter", "Population 1", "Population 2", "Definition")

giveparams$Definition <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Rate of sampling opinions or social interaction",
                    "Conservation cost",
                    "Rarity valuation",
                    "Strength of social influence (within population)",
                    "Fish movement (from opposite patch)",
                    "Strength of social influence (from opposite population)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "(ref:newparamtable) \\label{NewParamTable}")

```

(ref:influenceasymgraph) The difference in increasing outside social pressure on population 1 (the $\rho_1$ parameter is increased down the columns of graphs) versus increasing social pressure from population 1 onto population 2 (the $\rho_2$ parameter is increased across rows of graphs) which compares self-pressure to pressuring the other gropu.

```{r influenceAsymGraph, warning = FALSE, fig.cap = '(ref:influenceasymgraph) \\label{influenceAsymGraph}'}

plotList <- list()
k <- 0

rhoS <- c(0.01, 0.25, 0.5)
dS <- c(0.01, 0.25, 0.5)

timespan <- seq(from = 0, to = 150, by = 1)

for(i in 1:length(dS)){
  params_Default <- params_list
  params_Default[9] <- dS[i]
  #params_Default[16] <- dS[i]
  
  for(j in 1:length(rhoS)){
    k <- k+1
    #params_Default[9] <- rhoS[j]
    params_Default[18] <- rhoS[j]
  
   times <- timespan
   yStart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=yStart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
    if(k ==1){
      p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      scale_x_discrete(name = "", limits = c(0, 500, 950)) +
      scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
      # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
      # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
      scale_discrete_manual(values=cbbPalette, aesthetics = "colour", name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2]))) +
      ggtitle(paste("\u03C1", " 1 = ", params_Default[7], "\u03C1", " 2 = ", params_Default[16])) +
      theme(plot.title = element_text(size=7))
    plotList[[k]] <- p
    
    }else{
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        scale_x_discrete(name = "", limits = c(0, 500, 950)) +
        scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
        # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
        # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
        #theme(legend.position = "none") +
        ggtitle(paste("\u03C1", "1=", params_Default[9], "\u03C1", "2=", params_Default[18])) +
        theme(plot.title = element_text(size=7))
      plotList[[k]] <- p
    }
  }
}

yleft = richtext_grob("Population", rot=90)

bottom = richtext_grob(
  text = paste('Time (year)',  " F=", yStart[1], "X=", yStart[3])
)
  

  p_legend <- get_legend(plotList[[1]])

grid.arrange(arrangeGrob(plotList[[1]] + theme(legend.position="none"),
                         plotList[[2]] + theme(legend.position="none"),
                         plotList[[3]] + theme(legend.position="none"),
                         plotList[[4]] + theme(legend.position="none"),
                         plotList[[5]] + theme(legend.position="none"),
                         plotList[[6]] + theme(legend.position="none"),
                         plotList[[7]] + theme(legend.position="none"),
                         plotList[[8]] + theme(legend.position="none"),
                         plotList[[9]] + theme(legend.position="none"),
                         nrow=3, left = yleft, bottom = bottom),
             p_legend, 
             ncol=2,widths=c(2, 0.3))



```

(ref:influenceasym) Comparing the changing of $\rho_1$ and $\rho_2$ on overall dynamics of the asymmetrical case

```{r influenceAsym, warning = FALSE, fig.cap = '(ref:influenceasym) \\label{influenceAsym}', eval = FALSE}

params_list <- data.frame(r_1 = 0.40, s_1 = 0.8, h_1 = 0.25, k_1 = 1.014, w_1 = 0.2, c_1 = 1.5, d_1 = 0.5, m_1 = 0.2, rho_1 = 0.5,
                               r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 1.014, w_2 = 0.35, c_2 = 1.5, d_2 = 0.5, m_2 = 0.2, rho_2 = 0.1)

k <- 0

modeled_data <- data.frame(matrix(ncol=6,nrow=0, dimnames=list(NULL, c("Rho_1", "Rho_2", "F1_Final", "F2_Final", "X1_Final", "X2_Final"))))

rhos <- seq(from = 0, to = .5, by = .01)

timespan <- seq(from = 0, to = 200, by = 1)

for(i in 1:length(rhos)){
  params_Default <- params_list
  params_Default[9] <- rhos[i]
  
  for(j in 1:length(rhos)){
    k <- k+1
    #params_Default[9] <- rhoS[j]
    params_Default[18] <- rhos[j]
  
   times <- timespan
   yStart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=yStart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
    
    modeled_data[k,] <- c(rhos[i], rhos[j], out_Diff$F_1[max(out_Diff$time)], out_Diff$F_2[max(out_Diff$time)], out_Diff$X_1[max(out_Diff$time)], out_Diff$X_2[max(out_Diff$time)])
    
  }
}

write.csv(modeled_data, "rhoModel.csv")


```

(ref:rhocompare1) The effect of changing rho 1 and rho 2 on final dynamics of F1 (a) and F2 (b) using parameters from Table 2. Data was run for 200 years


```{r RhoCompare1, results = 'asis', echo = FALSE, fig.show = "hold", out.width = "50%", fig.cap = '(ref:rhocompare1) \\label{RhoCompare1}'}

#FOR SOME REASON THESE GRAPHS ARE NOT PROPERLY FILTERING 

#modeled_data <- read.csv("rhoModel.csv")

modeled_data <- read.csv("rhoModel.csv")

modeled_GraphDataF <- modeled_data %>%
  mutate(Status1 = case_when(
    F1_Final <= 0.1 ~ "StockCrash",
    F1_Final > 0.1 & F1_Final < 1 ~ "SomeFish",
    F1_Final > 0.99 ~ "AllFish"
    )) %>%
  mutate(Status2 = case_when(
    F2_Final <= 0.1 ~ "StockCrash",
    F2_Final > 0.1 & F2_Final < 1 ~ "SomeFish",
    F2_Final > 0.99 ~ "AllFish"
    ))

whereFish <- modeled_GraphDataF %>%
  filter(Status1 == "SomeFish" | Status2 == "SomeFish")


ggplot(modeled_GraphDataF, aes(x = Rho_2, y = Rho_1, fill = Status1)) + geom_tile(width=.05) +
  # scale_fill_gradient2(
  # low = "red",
  # mid = "green",
  # high = "purple",
  # #limits = c(1, max(growthRate$lambda_new)),
  # midpoint = 0.1,
  # space = "Lab",
  # na.value = "grey50") +
  scale_x_continuous(name = "Rho 2", expand=c(0,0)) + #, breaks = c(1, 3, 6, 9, 12)) +
  scale_y_continuous(name = "Rho 1", expand=c(0,0)) 
  # labs(fill = "Final Population") +
  # geom_segment(aes(x = 2.5, y = .075, xend = 2.5, yend = 0.105), size = .5, color = "#5A5A5A") +
  # geom_segment(aes(x = 11.5, y = .015, xend = 12.5, yend = 0.015), size = .5, color = "#5A5A5A") +
  # geom_step(data = paretospagheto , aes(x=Frequency, y=PercentChange), size = .5, color = "#5A5A5A") +
  # geom_text(x=8, y=.07, label="Population Increasing") +
  # geom_text(x=3.5, y=.015, label="Population Decreasing")

ggplot(modeled_GraphDataF, aes(x = Rho_2, y = Rho_1, fill = Status2)) + geom_tile(width=.05) +
  # scale_fill_gradient2(
  # low = "red",
  # mid = "green",
  # high = "purple",
  # #limits = c(1, max(growthRate$lambda_new)),
  # midpoint = 0.1,
  # space = "Lab",
  # na.value = "grey50") +
  scale_x_continuous(name = "Rho 2", expand=c(0,0)) + #, breaks = c(1, 3, 6, 9, 12)) +
  scale_y_continuous(name = "Rho 1", expand=c(0,0)) 
  # labs(fill = "Final Population") +
  # geom_segment(aes(x = 2.5, y = .075, xend = 2.5, yend = 0.105), size = .5, color = "#5A5A5A") +
  # geom_segment(aes(x = 11.5, y = .015, xend = 12.5, yend = 0.015), size = .5, color = "#5A5A5A") +
  # geom_step(data = paretospagheto , aes(x=Frequency, y=PercentChange), size = .5, color = "#5A5A5A") +
  # geom_text(x=8, y=.07, label="Population Increasing") +
  # geom_text(x=3.5, y=.015, label="Population Decreasing")
```

(ref:xcompare) The effect of changing rho 1 and rho 2 on final dynamics of X1 (a) and X2 (b) using parameters from Table 2. Data was run for 200 years

```{r XCompare, results = 'asis', echo = FALSE, fig.show = "hold", out.width = "50%", fig.cap = '(ref:xcompare) \\label{XCompare}'}

#modeled_data <- read.csv("rhoModel.csv")

modeled_data <- read.csv("rhoModel.csv")

modeled_GraphDataX <- modeled_data %>%
  mutate(Status1 = case_when(
    X1_Final <= 0.1 ~ "AllFishing",
    X1_Final > 0.1 & X1_Final < 1 ~ "SomeFishing",
    X1_Final == 1 ~ "NoneFishing"
    )) %>%
  mutate(Status2 = case_when(
    X2_Final <= 0.1 ~ "AllFishing",
    X2_Final > 0.1 & X2_Final < 1 ~ "NoneFishing", #All of these eventually go to NoneFishing
    X2_Final == 1 ~ "NoneFishing"
    ))

whyNA <- modeled_GraphDataX %>%
  filter(is.na(Status2))

modeled_GraphDataX$Status2 <- replace_na(modeled_GraphDataX$Status2, "NoneFishing")

# whereFishing <- modeled_GraphDataX %>%
#   filter(Status1 == "Some" | Status2 == "SomeFishing")
  

ggplot(modeled_GraphDataX, aes(x = Rho_2, y = Rho_1, fill = Status1)) + geom_tile(width=.05) +
  # scale_fill_gradient2(
  # low = "red",
  # mid = "green",
  # high = "purple",
  # #limits = c(1, max(growthRate$lambda_new)),
  # midpoint = 0.1,
  # space = "Lab",
  # na.value = "grey50") +
  scale_x_continuous(name = "Rho 2", expand=c(0,0)) + #, breaks = c(1, 3, 6, 9, 12)) +
  scale_y_continuous(name = "Rho 1", expand=c(0,0)) 
  # labs(fill = "Final Population") +
  # geom_segment(aes(x = 2.5, y = .075, xend = 2.5, yend = 0.105), size = .5, color = "#5A5A5A") +
  # geom_segment(aes(x = 11.5, y = .015, xend = 12.5, yend = 0.015), size = .5, color = "#5A5A5A") +
  # geom_step(data = paretospagheto , aes(x=Frequency, y=PercentChange), size = .5, color = "#5A5A5A") +
  # geom_text(x=8, y=.07, label="Population Increasing") +
  # geom_text(x=3.5, y=.015, label="Population Decreasing")

ggplot(modeled_GraphDataX, aes(x = Rho_2, y = Rho_1, fill = Status2)) + geom_tile(width=.05) +
  # scale_fill_gradient2(
  # low = "red",
  # mid = "green",
  # high = "purple",
  # #limits = c(1, max(growthRate$lambda_new)),
  # midpoint = 0.1,
  # space = "Lab",
  # na.value = "grey50") +
  scale_x_continuous(name = "Rho 2", expand=c(0,0)) + #, breaks = c(1, 3, 6, 9, 12)) +
  scale_y_continuous(name = "Rho 1", expand=c(0,0)) 
  # labs(fill = "Final Population") +
  # geom_segment(aes(x = 2.5, y = .075, xend = 2.5, yend = 0.105), size = .5, color = "#5A5A5A") +
  # geom_segment(aes(x = 11.5, y = .015, xend = 12.5, yend = 0.015), size = .5, color = "#5A5A5A") +
  # geom_step(data = paretospagheto , aes(x=Frequency, y=PercentChange), size = .5, color = "#5A5A5A") +
  # geom_text(x=8, y=.07, label="Population Increasing") +
  # geom_text(x=3.5, y=.015, label="Population Decreasing")
```

```{r whereFishingTest}

params_list <- data.frame(r_1 = 0.40, s_1 = 0.8, h_1 = 0.25, k_1 = 1.014, w_1 = 0.2, c_1 = 1.5, d_1 = 0.5, m_1 = 0.2, rho_1 = 0.5,
                               r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 1.014, w_2 = 0.35, c_2 = 1.5, d_2 = 0.5, m_2 = 0.2, rho_2 = 0.1)

timespan <- seq(from = 0, to = 1000, by = 1)


plotList <- list()


for(i in 1:length(whereFishing)){
  params_Default <- params_list
  params_Default[9] <- whereFishing$Rho_1[i]
  params_Default[18] <- whereFishing$Rho_2[i]
  

   times <- timespan
   yStart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=yStart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
    if(i ==1){
      p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      scale_x_discrete(name = "", limits = c(0, 500, 950)) +
      scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
      # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
      # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
      scale_discrete_manual(values=cbbPalette, aesthetics = "colour", name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2]))) +
      ggtitle(paste("\u03C1", " 1 = ", params_Default[9], "\u03C1", " 2 = ", params_Default[18])) +
      theme(plot.title = element_text(size=7))
    plotList[[i]] <- p
    
    }else{
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        scale_x_discrete(name = "", limits = c(0, 500, 950)) +
        scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
        # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
        # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
        #theme(legend.position = "none") +
        ggtitle(paste("\u03C1", "1=", params_Default[9], "\u03C1", "2=", params_Default[18])) +
        theme(plot.title = element_text(size=7))
      plotList[[i]] <- p
    }
}

yleft = richtext_grob("Population", rot=90)

bottom = richtext_grob(
  text = paste('Time (year)',  " F=", yStart[1], "X=", yStart[3])
)
  

  p_legend <- get_legend(plotList[[1]])

grid.arrange(arrangeGrob(plotList[[1]] + theme(legend.position="none"),
                         plotList[[2]] + theme(legend.position="none"),
                         plotList[[3]] + theme(legend.position="none"),
                         plotList[[4]] + theme(legend.position="none"),
                         plotList[[5]] + theme(legend.position="none"),
                         plotList[[6]] + theme(legend.position="none"),
                         plotList[[7]] + theme(legend.position="none"),
                         plotList[[8]] + theme(legend.position="none"),
                         plotList[[9]] + theme(legend.position="none"),
                         nrow=3, left = yleft, bottom = bottom),
             p_legend, 
             ncol=2,widths=c(2, 0.3))













```








