---
title: "Reveiwer1_RhoQ"
author: "Sophie Wulfing"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev="cairo_pdf", cache = TRUE)

library(deSolve)
library(tidyverse)
library(ggplot2)
library(gridExtra)
#library(kableExtra)
library(FME)
library(ggpubr)
library(gridtext)

#setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Soc_Ecol_Model")

params_dispersal <- data.frame(r_1 = 0.40, s_1 = 0.8, h_1 = 0.25, k_1 = 1.014, w_1 = 0.2, c_1 = 1.5, d_1 = 0.5, m_1 = 0.2, rho_1 = 0.5,
                               r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 1.014, w_2 = 0.35, c_2 = 1.5, d_2 = 0.5, m_2 = 0.2, rho_2 = 0.1)


cbbPalette <- c("#E69F00", "#009E73", "#567de9", "#D55E00") #F_1, F_2, X_1, X_2

# #WORKED - MINIMUM GROWTH RATE USED IN THAMPI
# params_list <- data.frame(r_1 = 0.16, s_1 = 0.8, h_1 = 0.5, k_1 = 0.17, w_1 = 1.44, c_1 = 0.6, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.16, s_2 = 0.8, h_2 = 0.5, k_2 = 0.17, w_2 = 1.44, c_2 = 0.6, d_2 = 0.3, m_2 = 0, rho_2 = 0)

#Extensions of Bauch Default but with my params I want changed. Goal: get this to oscillate
# params_list <- data.frame(r_1 = 0.35, s_1 = 0.8, h_1 = 0.5, k_1 = 0.17, w_1 = 1, c_1 = 1.68, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 0.17, w_2 = 1, c_2 = 1.68, d_2 = 0.3, m_2 = 0, rho_2 = 0)


#Extension of Bauch Default
# params_list <- data.frame(r_1 = 0.06, s_1 = 0.8, h_1 = 0.1, k_1 = 0.17, w_1 = 1, c_1 = 0.6, d_1 = 0.3, m_1 = 0, rho_1 = 0,
#                           r_2 = 0.06, s_2 = 0.8, h_2 = 0.1, k_2 = 0.17, w_2 = 1, c_2 = 0.6, d_2 = 0.3, m_2 = 0, rho_2 = 0)


xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)
timespan <- seq(from = 0, to = 1000, by = 1)

```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

```{r Bauch.Coupled, warning = FALSE}


##Maybe see what proportion of people can actually be fishing in this system to make it sustainable

Bauch.Coupled <- function (t, y, params) {
  ## first extract the state variables
  F_1 <- y[1] #Fish Pop var
  F_2 <- y[2] #Fish Pop var
  X_1 <- y[3] #conservationist var
  X_2 <- y[4] #conservationist var

  #TT <- y[5]
  
  ## now extract the parameters
  r_1 <- params["r_1"] # Net growth/fecundity
  s_1 <- params["s_1"] # supply and demand
  h_1 <- params["h_1"] # Harvesting efficiency
  k_1 <- params["k_1"] # Social learning rate
  w_1 <- params["w_1"] # conservation costs
  c_1 <- params["c_1"] # rarity valuation param
  d_1 <- params["d_1"] # social norm strength
  m_1 <- params["m_1"] # Inflow of fish from pop 2 to pop 1
  rho_1 <- params["rho_1"] # Influence of pop 2 on pop 1
  
  r_2 <- params["r_2"] # Net growth/fecundity
  s_2 <- params["s_2"] # supply and demand
  h_2 <- params["h_2"] # Harvesting efficiency
  k_2 <- params["k_2"] # Social learning rate
  w_2 <- params["w_2"] # conservation costs
  c_2 <- params["c_2"] # rarity valuation param
  d_2 <- params["d_2"] # social norm strength 
  m_2 <- params["m_2"] # Inflow of fish from pop 1 to pop 2
  rho_2 <- params["rho_2"] # Influence of pop 1 on pop 2
  
  ## now code the model equations
  dF_1dt <-  r_1 * F_1 * (1-F_1) - (h_1 * F_1 * (1-X_1))/(F_1+s_1) - m_2 * F_1 + m_1 * F_2
  dF_2dt <-  r_2 * F_2 * (1-F_2) - (h_2 * F_2 * (1-X_2))/(F_2+s_2) - m_1 * F_2 + m_2 * F_1
  
  dX_1dt <- k_1 * X_1 * (1-X_1) *(1/(F_1 + c_1) - w_1 + d_1*(2*X_1 - 1) + rho_1 * (2*X_2 - 1))
  dX_2dt <- k_2 * X_2 * (1-X_2) *(1/(F_2 + c_2) - w_2 + d_2*(2*X_2 - 1) + rho_2 * (2*X_1 - 1))
  
  ## combine results into a single vector
  dydt <- c(dF_1dt, dF_2dt, dX_1dt, dX_2dt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch2 <- params_dispersal

times <- timespan
xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_Bauch2
) %>%
  as.data.frame() -> out_Coupled

# out_Coupled %>%
#   gather(variable,value,-time) %>%
#   ggplot(aes(x=time,y=value,color=variable))+
#   geom_line(size=1)+
#   theme_classic()+
#   labs(x='Time (Year)',y='Population') +
#   theme(legend.title=element_blank()) +
#   scale_colour_discrete(labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2])))


```

(ref:influencerho2) The difference in increasing social pressure within population 1 (the $d_1$ parameter is increased down the columns of graphs) versus increasing social pressure from population 1 onto population 2 (the $\rho_2$ parameter is increased across rows of graphs) which models increasing the social pressure of the sustainable population onto the unsustainable. 

```{r influencerho2, warning = FALSE, fig.cap = '(ref:influencerho2) \\label{influenceRho2}'}

rhoparams <- c(0, 0.25, 0.5)
dparams <- c(0, 0.25, 0.5)

plotList <- list()
k <- 0

for(i in 1:length(dparams)){
  params_Default <- params_list
  params_Default[7] <- dparams[i]
  #params_Default[16] <- dparams[i]
  
  for(j in 1:length(rhoparams)){
    k <- k+1
    #params_Default[9] <- rhoparams[j]
    params_Default[18] <- rhoparams[j]
  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
    if(k ==1){
      p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      scale_x_discrete(name = "", limits = c(0, 500, 950)) +
      scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
      # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
      # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
      scale_discrete_manual(values=cbbPalette, aesthetics = "colour", name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2]))) +
      ggtitle(paste("\u03B4 = ", params_Default[7], "\u03C1 = ", params_Default[18])) +
      theme(plot.title = element_text(size=7))
    plotList[[k]] <- p
    
    }else{
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        scale_x_discrete(name = "", limits = c(0, 500, 950)) +
        scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
        # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
        # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
        #theme(legend.position = "none") +
        ggtitle(paste("\u03B4 = ", params_Default[7], "\u03C1 = ", params_Default[18])) +
        theme(plot.title = element_text(size=7))
      plotList[[k]] <- p
    }
  }
}

yleft = richtext_grob("Population", rot=90)

bottom = richtext_grob(
  text = 'Time (year)'
)
  

  p_legend <- get_legend(plotList[[1]])

grid.arrange(arrangeGrob(plotList[[1]] + theme(legend.position="none"),
                         plotList[[2]] + theme(legend.position="none"),
                         plotList[[3]] + theme(legend.position="none"),
                         plotList[[4]] + theme(legend.position="none"),
                         plotList[[5]] + theme(legend.position="none"),
                         plotList[[6]] + theme(legend.position="none"),
                         plotList[[7]] + theme(legend.position="none"),
                         plotList[[8]] + theme(legend.position="none"),
                         plotList[[9]] + theme(legend.position="none"),
                         nrow=3, left = yleft, bottom = bottom),
             p_legend, 
             ncol=2,widths=c(2, 0.3))

```

(ref:influencerho1) The difference in increasing social pressure within population 2 (the $d_2$ parameter is increased down the columns of graphs) versus increasing social pressure from population 2 onto population 1 (the $\rho_1$ parameter is increased across rows of graphs) which models increasing the social pressure of the unsustainable population on the sustainable. 

```{r influencerho1, warning = FALSE, fig.cap = '(ref:influencerho1) \\label{influenceRho1}'}

rhoparams <- c(0, 0.25, 0.5)
dparams <- c(0, 0.25, 0.5)

plotList <- list()
k <- 0

for(i in 1:length(dparams)){
  params_Default <- params_list
  #params_Default[7] <- dparams[i]
  params_Default[16] <- dparams[i]
  
  for(j in 1:length(rhoparams)){
    k <- k+1
    params_Default[9] <- rhoparams[j]
    #params_Default[18] <- rhoparams[j]
  
   times <- timespan
    xstart <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

    ode(
      func=Bauch.Coupled,
      y=xstart,
      times=times,
      parms=params_Default
    ) %>%
      as.data.frame() -> out_Diff
    
    if(k ==1){
      p <- out_Diff %>%
      gather(variable,value,-time) %>%
      ggplot(aes(x=time,y=value,color=variable))+
      geom_line(size=1)+
      theme_classic()+
      scale_colour_manual(values=cbbPalette) +
      scale_x_discrete(name = "", limits = c(0, 500, 950)) +
      scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
      # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
      # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
      scale_discrete_manual(values=cbbPalette, aesthetics = "colour", name = "Variable", labels=c(expression('F'[1]), expression('F'[2]), expression('X'[1]), expression('X'[2]))) +
      ggtitle(paste("\u03B4 = ", params_Default[7], "\u03C1 = ", params_Default[18])) +
      theme(plot.title = element_text(size=7))
    plotList[[k]] <- p
    
    }else{
      p <- out_Diff %>%
        gather(variable,value,-time) %>%
        ggplot(aes(x=time,y=value,color=variable))+
        geom_line(size=1)+
        theme_classic()+
        scale_colour_manual(values=cbbPalette) +
        scale_x_discrete(name = "", limits = c(0, 500, 950)) +
        scale_y_discrete(name = "", limits = c(0, 0.5, 0.95)) +
        # scale_x_discrete(name = "Time (Year)", limits = c(0,750)) +
        # scale_y_discrete(name = "Population", limits = c(0, 0.5, 1)) +
        #theme(legend.position = "none") +
        ggtitle(paste("\u03B4 = ", params_Default[7], "\u03C1 = ", params_Default[18])) +
        theme(plot.title = element_text(size=7))
      plotList[[k]] <- p
    }
  }
}

yleft = richtext_grob("Population", rot=90)

bottom = richtext_grob(
  text = 'Time (year)'
)
  

  p_legend <- get_legend(plotList[[1]])

grid.arrange(arrangeGrob(plotList[[1]] + theme(legend.position="none"),
                         plotList[[2]] + theme(legend.position="none"),
                         plotList[[3]] + theme(legend.position="none"),
                         plotList[[4]] + theme(legend.position="none"),
                         plotList[[5]] + theme(legend.position="none"),
                         plotList[[6]] + theme(legend.position="none"),
                         plotList[[7]] + theme(legend.position="none"),
                         plotList[[8]] + theme(legend.position="none"),
                         plotList[[9]] + theme(legend.position="none"),
                         nrow=3, left = yleft, bottom = bottom),
             p_legend, 
             ncol=2,widths=c(2, 0.3))

```


```{r rhoExplore}

param_seq <- seq(0, .5, by = .01)
finalDynamics <- data.frame(matrix(, nrow = (length(param_seq)), ncol = 3))
colnames(finalDynamics) <- c("Parameter_New", "F1", "F2")
timespan <- seq(from = 0, to = 100, by = .1)

starting_defaultparams <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)

for(i in 1:length(param_seq)){
  #Run momdel with each new param combo
  
  # params_test <- params_dispersal  
  # params_test$m_1 <- param_seq[i]
  # params_test$m_2 <- param_seq[i]
  
  
  params_test <- params_dispersal
  params_test$rho_1 <- param_seq[i]
  params_test$rho_2 <- param_seq[i]
  
  finalDynamics$Parameter_New[i] <- param_seq[i]

  ode(
    func=Bauch.Coupled,
    y=starting_defaultparams,
    times=timespan,
    parms=params_test
  ) %>%
    as.data.frame() -> out_Coupled

  #Compile output into Dataframe
  finalDynamics$F1[i] <- out_Coupled[nrow(out_Coupled), 2]
  finalDynamics$F2[i] <- out_Coupled[nrow(out_Coupled), 3]

}

param_seq <- seq(0, .5, by = .01)
finalDynamics1 <- data.frame(matrix(, nrow = (length(param_seq)), ncol = 3))
colnames(finalDynamics1) <- c("Parameter_New", "F1", "F2")

for(i in 1:length(param_seq)){
  #Run momdel with each new param combo
  
  # params_test <- params_dispersal  
  # params_test$m_1 <- param_seq[i]
  # params_test$m_2 <- param_seq[i]
  
  
  params_test <- params_dispersal
  params_test$rho_1 <- param_seq[i]
  #params_test$rho_2 <- param_seq[i]
  
  finalDynamics1$Parameter_New[i] <- param_seq[i]

  ode(
    func=Bauch.Coupled,
    y=starting_defaultparams,
    times=timespan,
    parms=params_test
  ) %>%
    as.data.frame() -> out_Coupled

  #Compile output into Dataframe
  finalDynamics1$F1[i] <- out_Coupled[nrow(out_Coupled), 2]
  finalDynamics1$F2[i] <- out_Coupled[nrow(out_Coupled), 3]

}

param_seq <- seq(0, .5, by = .01)
finalDynamics2 <- data.frame(matrix(, nrow = (length(param_seq)), ncol = 3))
colnames(finalDynamics2) <- c("Parameter_New", "F1", "F2")

for(i in 1:length(param_seq)){
  #Run momdel with each new param combo
  
  # params_test <- params_dispersal  
  # params_test$m_1 <- param_seq[i]
  # params_test$m_2 <- param_seq[i]
  
  
  params_test <- params_dispersal
  #params_test$rho_1 <- param_seq[i]
  params_test$rho_2 <- param_seq[i]
  
  finalDynamics2$Parameter_New[i] <- param_seq[i]

  ode(
    func=Bauch.Coupled,
    y=starting_defaultparams,
    times=timespan,
    parms=params_test
  ) %>%
    as.data.frame() -> out_Coupled

  #Compile output into Dataframe
  finalDynamics2$F1[i] <- out_Coupled[nrow(out_Coupled), 2]
  finalDynamics2$F2[i] <- out_Coupled[nrow(out_Coupled), 3]

}

```

(ref:rhoexploregrapheach) Each rho individually. Ok so here's my confusion, above in \ref{influenceAsym} I say that incorporating new information will increase stability but here, as pop 2 (which is unsustainable) listens to pop 1 more, the whole thing crashes. Earlier we said this was because pop 1 is continuing to fish, so therefore encouraging pop 2 to fish more (looking at graph a)

```{r rhoExploreGraphEach, results = 'asis', echo = FALSE, fig.show = "hold", out.width = "50%", fig.cap = '(ref:rhoexploregrapheach) \\label{rhoExploreGraphEach}'}

patchcolors <- c("#567de9", "#D55E00")

finalDynamics1_long <- gather(finalDynamics1, plot, Population, F1:F2, factor_key=TRUE)

finalDynamics1_long %>%
  ggplot(aes(Parameter_New, Population, color=plot)) +
  geom_line() +
  geom_point(alpha=0.5, size=1) +
  scale_color_manual(values = patchcolors) +
  ggtitle('a)') +
  labs(y="Final Fish Population", x="rho1 Value") +
  theme_classic() +  
  theme(legend.position = "none")

#Scatter plot

patchcolors <- c("#567de9", "#D55E00")

finalDynamics2_long <- gather(finalDynamics2, plot, Population, F1:F2, factor_key=TRUE)

finalDynamics2_long %>%
  ggplot(aes(Parameter_New, Population, color=plot)) +
  geom_line() +
  geom_point(alpha=0.5, size=1) +
  ggtitle('b)') +
  labs(y="Final Fish Population", x="rho2 Value") +
  theme_classic() +
  theme(legend.title=element_blank()) +
  scale_colour_manual(values = patchcolors, labels=c(expression('F'[1]), expression('F'[2])))




```

(ref:rhoexploregraph) Final fish populations after 100 years in the two-patch fishing model where the $F_1$ population in patch 1 is fished sustainably but human population 1 has a lower social influence than humans in patch 2, where $F_2$ is being fished unsustainably. Both $\rho_1$ and $\rho_2$ were increased simultaneously.

```{r rhoExploreGraph, fig.cap = '(ref:rhoexploregraph) \\label{rhoExploreGraph}'}

patchcolors <- c("#567de9", "#D55E00")

finalDynamics_long <- gather(finalDynamics, plot, Population, F1:F2, factor_key=TRUE)

finalDynamics_long %>%
  ggplot(aes(Parameter_New, Population, color=plot)) +
  geom_line() +
  geom_point(alpha=0.5, size=1) +
  scale_color_manual(values = patchcolors) +
  labs(y="Final Fish Population", x="\u03C1 Value") +
  theme_classic() +
  theme(legend.title=element_blank()) +
  scale_colour_manual(values = patchcolors, labels=c(expression('F'[1]), expression('F'[2])))




```











