---
title: "SocModelScenarios"
author: "Sophie Wulfing"
date: "2023-03-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, warning = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Soc_Ecol_Model")

library(deSolve)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(kableExtra)
#library(ODEsensitivity)
library(FME)

param_names <- c("r_1", "s_1", "h_1", "k_1", "w_1", "c_1", "d_1", "i_1", "rho_1",
                 "r_2", "s_2", "h_2", "k_2", "w_2", "c_2", "d_2", "i_2", "rho_2")

params_default <- data.frame(r_1 = 0.35, s_1 = 0.8, h_1 = 0.5, k_1 = 1.014, w_1 = 0.35, c_1 = 1.5, d_1 = 0.5, i_1 = 0.2, rho_1 = 0.5,
                             r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 1.014, w_2 = 0.35, c_2 = 1.5, d_2 = 0.5, i_2 = 0.2, rho_2 = 0.5)

starting_defaultparams <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)
timespan <- seq(from = 0, to = 100, by = .1)

```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

Function:

$\frac{dF_1}{dt} = r_1F_1(1-F_1)-\frac{h_1*F_1(1-X_1)}{F_1 + s_1} -i_2F_1 + i_1F_2$

$\frac{dF_2}{dt} = r_2F_2(1-F_2)-\frac{h_2*F_2(1-X_2)}{F_2 + s_2} -i_1F_2 + i_2F_1$

$\frac{dX_1}{dt} =  k_1X_1(1-X_1) [\frac{1}{F_1+c_1} - \omega_1 + d_1(2X_1 - 1) + \rho_1(2X_2 - 1)]$

$\frac{dX_2}{dt} = k_2X_2(1-X_2)  [\frac{1}{F_2+c_2} - \omega_2 + d_2(2X_2 - 1) +  \rho_2(2X_1 - 1)]$

```{r defaultparamtable, results = 'asis', echo = FALSE}

giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Def")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "i", "$\\rho$")
giveparams$Population_1 <- c(params_default[1:9])
giveparams$Population_2 <- c(params_default[10:18])

giveparams$Def <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Social learning rate",
                    "Conservation cost",
                    "Rarity valuation",
                    "Social norm strength (within pop)",
                    "Fish immigration (from patch)",
                    "Social norm strength (opposite pop)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "Default parameter values used in this analysis")

```

```{r startingtable, results = 'asis', echo = FALSE}
givestart <- data.frame(matrix(, nrow = 2, ncol = 3))
colnames(givestart) <- c("Parameter", "Population_1", "Population_2")
givestart$Parameter <- c("F", "X")
givestart$Population_1 <- c(starting_defaultparams[1], starting_defaultparams[3])
givestart$Population_2 <- c(starting_defaultparams[2], starting_defaultparams[4])


knitr::kable(givestart, format = "latex", booktabs = TRUE, caption = "Starting values used in this analysis")


```

\newpage

```{r ModelNoSocial, fig.cap = "Model without social dynamics", warning = FALSE}
ModelNoSocial <- function (t, y, params) {
  ## first extract the state variables
  F_1 <- y[1] #Fish Pop var
  F_2 <- y[2] #Fish Pop var

  #TT <- y[5]
  
  ## now extract the parameters
  r_1 <- params["r_1"] # Net growth/fecundity
  s_1 <- params["s_1"] # supply and demand
  h_1 <- params["h_1"] # Harvesting efficiency
 
  r_2 <- params["r_2"] # Net growth/fecundity
  s_2 <- params["s_2"] # supply and demand
  h_2 <- params["h_2"] # Harvesting efficiency
  
  i_1 <- params["i_1"] # Inflow of fish from pop 2 to pop 1
  i_2 <- params["i_2"] # Inflow of fish from pop 1 to pop 2

  ## now code the model equations
  dF_1dt <-  r_1 * F_1 * (1-F_1) - (h_1 * F_1 * .76)/(F_1+s_1) - i_2 * F_1 + i_1 * F_2
  dF_2dt <-  r_2 * F_2 * (1-F_2) - (h_2 * F_2 * .26)/(F_2+s_2) - i_1 * F_2 + i_2 * F_1

  
  ## combine results into a single vector
  dydt <- c(dF_1dt, dF_2dt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch2 <- data.frame(r_1 = params_default$r_1, s_1 = params_default$s_1, h_1 = params_default$h_1, i_1 = params_default$i_1,
                            r_2 = params_default$r_2, s_2 = params_default$s_2, h_2 = params_default$h_2, i_2 = params_default$i_2)

times <- timespan
xstart <- starting_defaultparams[1:2] # See supplemental data for other starting params

ode(
  func=ModelNoSocial,
  y=xstart,
  times=times,
  parms=params_Bauch2
) %>%
  as.data.frame() -> out_Coupled

out_Coupled %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

```

Observations: 

* Fishing remains sustainable UNLESS more than 50% of people are fishing

\newpage

```{r NewModel, fig.cap = "New Model with social dynamics", warning = FALSE}


##Maybe see what proportion of people can actually be fishing in this system to make it sustainable

Bauch.Coupled <- function (t, y, params) {
  ## first extract the state variables
  F_1 <- y[1] #Fish Pop var
  F_2 <- y[2] #Fish Pop var
  X_1 <- y[3] #conservationist var
  X_2 <- y[4] #conservationist var

  #TT <- y[5]
  
  ## now extract the parameters
  r_1 <- params["r_1"] # Net growth/fecundity
  s_1 <- params["s_1"] # supply and demand
  h_1 <- params["h_1"] # Harvesting efficiency
  k_1 <- params["k_1"] # Social learning rate
  w_1 <- params["w_1"] # conservation costs
  c_1 <- params["c_1"] # rarity valuation param
  d_1 <- params["d_1"] # social norm strength
  
  r_2 <- params["r_2"] # Net growth/fecundity
  s_2 <- params["s_2"] # supply and demand
  h_2 <- params["h_2"] # Harvesting efficiency
  k_2 <- params["k_2"] # Social learning rate
  w_2 <- params["w_2"] # conservation costs
  c_2 <- params["c_2"] # rarity valuation param
  d_2 <- params["d_2"] # social norm strength 
  
  i_1 <- params["i_1"] # Inflow of fish from pop 2 to pop 1
  i_2 <- params["i_2"] # Inflow of fish from pop 1 to pop 2
  rho_1 <- params["rho_1"] # Influence of pop 2 on pop 1
  rho_2 <- params["rho_2"] # Influence of pop 1 on pop 2
  
  ## now code the model equations
  dF_1dt <-  r_1 * F_1 * (1-F_1) - (h_1 * F_1 * (1-X_1))/(F_1+s_1) - i_2 * F_1 + i_1 * F_2
  dF_2dt <-  r_2 * F_2 * (1-F_2) - (h_2 * F_2 * (1-X_2))/(F_2+s_2) - i_1 * F_2 + i_2 * F_1
  
  dX_1dt <- k_1 * X_1 * (1-X_1) *(1/(F_1 + c_1) - w_1 + d_1*(2*X_1 - 1) + rho_1 * (2*X_2 - 1))
  dX_2dt <- k_2 * X_2 * (1-X_2) *(1/(F_2 + c_2) - w_2 + d_2*(2*X_2 - 1) + rho_2 * (2*X_1 - 1))
  
  ## combine results into a single vector
  dydt <- c(dF_1dt, dF_2dt, dX_1dt, dX_2dt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch2 <- params_default

times <- timespan
xstart <- starting_defaultparams # See supplemental data for other starting params

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_Bauch2
) %>%
  as.data.frame() -> out_Coupled

out_Coupled %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

```

Observations:

* Still decreasing. Indicates that human dynamics consistently have fishers above 50%

\newpage

# Scenarios

## One group tanking whole system

```{r socparamtable, results = 'asis', echo = FALSE}

params_social <- params_default
params_social$r_1 <- 0.4
params_social$r_2 <- 0.4
params_social$w_1 <- 0.2
params_social$w_2 <- 0.2
params_social$h_1 <- 0.25
params_social$h_2 <- 0.25

# params_social$c_1 <- 0.25
# params_social$c_2 <- 0.25


giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Def")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "i", "$\\rho$")
giveparams$Population_1 <- c(params_social[1:9])
giveparams$Population_2 <- c(params_social[10:18])

giveparams$Def <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Social learning rate",
                    "Conservation cost",
                    "Rarity valuation",
                    "Social norm strength (within pop)",
                    "Fish immigration (from patch)",
                    "Social norm strength (opposite pop)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "Default parameter values used in this analysis")

```

```{r socialscenario, fig.cap = "Changing fish growth, conservaiton cost, and harvesting efficiency for sustainable practices"}


ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_social
) %>%
  as.data.frame() -> out_social

out_social %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

#Social inequity scenario where rho_1 = 0.25 and h_1 = 1

```

```{r socialscenario_unsustainable, fig.cap = "One group unsustainable practices scenario. Shows that one groups bad fishing can tank whole system"}

params_social$r_1 <- 0.4
params_social$r_2 <- 0.35
params_social$w_1 <- 0.2
params_social$w_2 <- 0.35
params_social$h_1 <- 0.25
params_social$h_2 <- 0.5

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_social
) %>%
  as.data.frame() -> out_social

out_social %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

#Social inequity scenario where rho_1 = 0.25 and h_1 = 1

```

```{r socialscenario_revive, fig.cap = "However, changing the rarity valuation parameters can recover the system"}

params_social$c_1 <- 0.25
params_social$c_2 <- 0.25

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_social
) %>%
  as.data.frame() -> out_social

out_social %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

#Social inequity scenario where rho_1 = 0.25 and h_1 = 1

```

\newpage

## Heirarchy in decision making

```{r heirparamtable, results = 'asis', echo = FALSE}

params_heir <- params_default
params_heir$h_2 <- 0.25
params_heir$k_1 <- 0.17
params_heir$i_1 <- 0.1
params_heir$i_2 <- 0.4
#params_heir$rho_1 <- .1


# params_social$c_1 <- 0.25
# params_social$c_2 <- 0.25


giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Def")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "i", "$\\rho$")
giveparams$Population_1 <- c(params_heir[1:9])
giveparams$Population_2 <- c(params_heir[10:18])

giveparams$Def <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Social learning rate",
                    "Conservation cost",
                    "Rarity valuation",
                    "Social norm strength (within pop)",
                    "Fish immigration (from patch)",
                    "Social norm strength (opposite pop)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "Default parameter values used in this analysis")

```

```{r heirscenario, fig.cap = "model with different fishing conditions in each patch"}

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=seq(from = 0, to = 200, by = .1),
  parms=params_heir
) %>%
  as.data.frame() -> out_heir

out_heir %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')
```

```{r heirscenario_inequity, fig.cap = "model with social inequity. idk does similar stuff but allows x1 to fish more"}

params_heir$rho_1 <- .1

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=seq(from = 0, to = 200, by = .1),
  parms=params_heir
) %>%
  as.data.frame() -> out_heir

out_heir %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')
```

\newpage

## Dispersion Exploration

```{r dispersionparamtable, results = 'asis', echo = FALSE}

params_dispersion <- params_default
params_dispersion$r_1 <- 0.4
#params_dispersion$r_2 <- 0.4
params_dispersion$w_1 <- 0.2
#params_dispersion$w_2 <- 0.2
params_dispersion$h_1 <- 0.25
#params_dispersion$h_2 <- 0.25
params_dispersion$i_1 <- 0
params_dispersion$i_2 <- 0

giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Def")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "i", "$\\rho$")
giveparams$Population_1 <- c(params_dispersion[1:9])
giveparams$Population_2 <- c(params_dispersion[10:18])

giveparams$Def <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Social learning rate",
                    "Conservation cost",
                    "Rarity valuation",
                    "Social norm strength (within pop)",
                    "Fish immigration (from patch)",
                    "Social norm strength (opposite pop)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "Default parameter values used in this analysis")

```

```{r dispersionscenario, fig.cap = "no dispersion. Unsustainable practices in one patch"}


ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_dispersion
) %>%
  as.data.frame() -> out_dispersion

out_dispersion %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

#Social inequity scenario where rho_1 = 0.25 and h_1 = 1

```

```{r dispersionscenario_slow, fig.cap = "slow dispersion. Unsustainable practices in one patch. Note: no adjustment to rho or d could fix fishing scenario"}

params_dispersion$i_1 <- 0.1
params_dispersion$i_2 <- 0.1


ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_dispersion
) %>%
  as.data.frame() -> out_dispersion

out_dispersion %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

#Social inequity scenario where rho_1 = 0.25 and h_1 = 1

```


```{r dispersionscenario_norho, fig.cap = "No social dynamics. Taking away rho doesn't really improve dynamics, but instead allows pop1 to take advantage of their sustainable fishing"}
params_dispersion <- params_default
params_dispersion$r_1 <- 0.4
#params_dispersion$r_2 <- 0.4
params_dispersion$w_1 <- 0.2
#params_dispersion$w_2 <- 0.2
params_dispersion$h_1 <- 0.25
#params_dispersion$h_2 <- 0.25
params_dispersion$rho_1 <- 0
params_dispersion$rho_2 <- 0
params_dispersion$i_1 <- 0
params_dispersion$i_2 <- 0


ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_dispersion
) %>%
  as.data.frame() -> out_dispersion

out_dispersion %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

#Social inequity scenario where rho_1 = 0.25 and h_1 = 1

```

\newpage

```{r Madagascarparamtable, results = 'asis', echo = FALSE}

params_Madagascar <- params_default
params_Madagascar$r_1 <- 0.07 #Max growth rate calcualted from our matrix with no fishing and 10% change in survival
params_Madagascar$r_2 <- 0.07

params_Madagascar$h_1 <- 0.1
params_Madagascar$h_2 <- 0.1

params_Madagascar$i_1 <- 0.07 # Calculated from our stable stage dist.

giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Def")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "i", "$\\rho$")
giveparams$Population_1 <- c(params_Madagascar[1:9])
giveparams$Population_2 <- c(params_Madagascar[10:18])

giveparams$Def <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Social learning rate",
                    "Conservation cost",
                    "Rarity valuation",
                    "Social norm strength (within pop)",
                    "Fish immigration (from patch)",
                    "Social norm strength (opposite pop)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "Default parameter values used in this analysis")

```

```{r Madagascarscenario, fig.cap = "Hypothetical Madagascar. Can change H, S, omega, rho"}


ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_Madagascar
) %>%
  as.data.frame() -> out_Madagascar

out_Madagascar %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

#Social inequity scenario where rho_1 = 0.25 and h_1 = 1

```

```{r Madagascarscenari_Inequity, fig.cap = "Hypothetical Madagascar with social inequity"}

params_Madagascar$rho_1 <- .1


ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_Madagascar
) %>%
  as.data.frame() -> out_Madagascar

out_Madagascar %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

#Social inequity scenario where rho_1 = 0.25 and h_1 = 1

```