---
title: "SocModelAnalysis"
author: "Sophie Wulfing"
date: "2023-03-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, warning = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Soc_Ecol_Model")

library(deSolve)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(kableExtra)
#library(ODEsensitivity)
library(FME)

param_names <- c("r_1", "s_1", "h_1", "k_1", "w_1", "c_1", "d_1", "i_1", "rho_1",
                 "r_2", "s_2", "h_2", "k_2", "w_2", "c_2", "d_2", "i_2", "rho_2")

params_default <- data.frame(r_1 = 0.35, s_1 = 0.8, h_1 = 0.5, k_1 = 1.014, w_1 = 0.35, c_1 = 1.5, d_1 = 0.5, i_1 = 0.2, rho_1 = 0.5,
                             r_2 = 0.35, s_2 = 0.8, h_2 = 0.5, k_2 = 1.014, w_2 = 0.35, c_2 = 1.5, d_2 = 0.5, i_2 = 0.2, rho_2 = 0.5)

starting_defaultparams <- c(F_1 = 0.406, F_2 = 0.406, X_1 = .24, X_2 = .24)
timespan <- seq(from = 0, to = 100, by = .1)

```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

Function:

$\frac{dF_1}{dt} = r_1F_1(1-F_1)-\frac{h_1*F_1(1-X_1)}{F_1 + s_1} -i_2F_1 + i_1F_2$

$\frac{dF_2}{dt} = r_2F_2(1-F_2)-\frac{h_2*F_2(1-X_2)}{F_2 + s_2} -i_1F_2 + i_2F_1$

$\frac{dX_1}{dt} =  k_1X_1(1-X_1) [\frac{1}{F_1+c_1} - \omega_1 + d_1(2X_1 - 1) + \rho_1(2X_2 - 1)]$

$\frac{dX_2}{dt} = k_2X_2(1-X_2)  [\frac{1}{F_2+c_2} - \omega_2 + d_2(2X_2 - 1) +  \rho_2(2X_1 - 1)]$

```{r defaultparamtable, results = 'asis', echo = FALSE}

giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Def")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "i", "$\\rho$")
giveparams$Population_1 <- c(params_default[1:9])
giveparams$Population_2 <- c(params_default[10:18])

giveparams$Def <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Social learning rate",
                    "Conservation cost",
                    "Rarity valuation",
                    "Social norm strength (within pop)",
                    "Fish immigration (from patch)",
                    "Social norm strength (opposite pop)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "Default parameter values used in this analysis")

```

```{r startingtable, results = 'asis', echo = FALSE}
givestart <- data.frame(matrix(, nrow = 2, ncol = 3))
colnames(givestart) <- c("Parameter", "Population_1", "Population_2")
givestart$Parameter <- c("F", "X")
givestart$Population_1 <- c(starting_defaultparams[1], starting_defaultparams[3])
givestart$Population_2 <- c(starting_defaultparams[2], starting_defaultparams[4])


knitr::kable(givestart, format = "latex", booktabs = TRUE, caption = "Starting values used in this analysis")


```

\newpage

```{r ModelNoSocial, fig.cap = "Model without social dynamics", warning = FALSE}
ModelNoSocial <- function (t, y, params) {
  ## first extract the state variables
  F_1 <- y[1] #Fish Pop var
  F_2 <- y[2] #Fish Pop var

  #TT <- y[5]
  
  ## now extract the parameters
  r_1 <- params["r_1"] # Net growth/fecundity
  s_1 <- params["s_1"] # supply and demand
  h_1 <- params["h_1"] # Harvesting efficiency
 
  r_2 <- params["r_2"] # Net growth/fecundity
  s_2 <- params["s_2"] # supply and demand
  h_2 <- params["h_2"] # Harvesting efficiency
  
  i_1 <- params["i_1"] # Inflow of fish from pop 2 to pop 1
  i_2 <- params["i_2"] # Inflow of fish from pop 1 to pop 2

  ## now code the model equations
  dF_1dt <-  r_1 * F_1 * (1-F_1) - (h_1 * F_1 * 0.50)/(F_1+s_1) - i_2 * F_1 + i_1 * F_2
  dF_2dt <-  r_2 * F_2 * (1-F_2) - (h_2 * F_2 * 0.50)/(F_2+s_2) - i_1 * F_2 + i_2 * F_1

  
  ## combine results into a single vector
  dydt <- c(dF_1dt, dF_2dt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch2 <- data.frame(r_1 = params_default$r_1, s_1 = params_default$s_1, h_1 = params_default$h_1, i_1 = params_default$i_1,
                            r_2 = params_default$r_2, s_2 = params_default$s_2, h_2 = params_default$h_2, i_2 = params_default$i_2)

times <- timespan
xstart <- starting_defaultparams[1:2] # See supplemental data for other starting params

ode(
  func=ModelNoSocial,
  y=xstart,
  times=times,
  parms=params_Bauch2
) %>%
  as.data.frame() -> out_Coupled

out_Coupled %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

```

Observations: 

* Fishing remains sustainable UNLESS more than 50% of people are fishing

\newpage

```{r NewModel, fig.cap = "New Model with social dynamics", warning = FALSE}


##Maybe see what proportion of people can actually be fishing in this system to make it sustainable

Bauch.Coupled <- function (t, y, params) {
  ## first extract the state variables
  F_1 <- y[1] #Fish Pop var
  F_2 <- y[2] #Fish Pop var
  X_1 <- y[3] #conservationist var
  X_2 <- y[4] #conservationist var

  #TT <- y[5]
  
  ## now extract the parameters
  r_1 <- params["r_1"] # Net growth/fecundity
  s_1 <- params["s_1"] # supply and demand
  h_1 <- params["h_1"] # Harvesting efficiency
  k_1 <- params["k_1"] # Social learning rate
  w_1 <- params["w_1"] # conservation costs
  c_1 <- params["c_1"] # rarity valuation param
  d_1 <- params["d_1"] # social norm strength
  
  r_2 <- params["r_2"] # Net growth/fecundity
  s_2 <- params["s_2"] # supply and demand
  h_2 <- params["h_2"] # Harvesting efficiency
  k_2 <- params["k_2"] # Social learning rate
  w_2 <- params["w_2"] # conservation costs
  c_2 <- params["c_2"] # rarity valuation param
  d_2 <- params["d_2"] # social norm strength 
  
  i_1 <- params["i_1"] # Inflow of fish from pop 2 to pop 1
  i_2 <- params["i_2"] # Inflow of fish from pop 1 to pop 2
  rho_1 <- params["rho_1"] # Influence of pop 2 on pop 1
  rho_2 <- params["rho_2"] # Influence of pop 1 on pop 2
  
  ## now code the model equations
  dF_1dt <-  r_1 * F_1 * (1-F_1) - (h_1 * F_1 * (1-X_1))/(F_1+s_1) - i_2 * F_1 + i_1 * F_2
  dF_2dt <-  r_2 * F_2 * (1-F_2) - (h_2 * F_2 * (1-X_2))/(F_2+s_2) - i_1 * F_2 + i_2 * F_1
  
  dX_1dt <- k_1 * X_1 * (1-X_1) *(1/(F_1 + c_1) - w_1 + d_1*(2*X_1 - 1) + rho_1 * (2*X_2 - 1))
  dX_2dt <- k_2 * X_2 * (1-X_2) *(1/(F_2 + c_2) - w_2 + d_2*(2*X_2 - 1) + rho_2 * (2*X_1 - 1))
  
  ## combine results into a single vector
  dydt <- c(dF_1dt, dF_2dt, dX_1dt, dX_2dt)
  ## return result as a list!
  
  list(dydt)
}

params_Bauch2 <- params_default

times <- timespan
xstart <- starting_defaultparams # See supplemental data for other starting params

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_Bauch2
) %>%
  as.data.frame() -> out_Coupled

out_Coupled %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

```

Observations:

* Still decreasing


\newpage

```{r BifurcationPlots}

#Ok I have no clue wtf this is telling me

n <- 100 # NUMBER OF SIMULATIONS
param.name <- "w_1" # PARAM TO PERTURB
param.seq <- seq(0,1,length = 50) # RANGE OF PARAMETER

Pars <- params_default
Time <- seq(0, 10, length = n)
State <- starting_defaultparams

param.index <- which(param.name == names(Pars))
out <- list()
for (i in 1:length(param.seq)){
  out[[i]] <- matrix(0, n, length(State))}

for (i in 1:length(param.seq)) {
  # set params
  Pars.loop <- Pars
  Pars.loop[param.index] <- param.seq[i]
  # converge
  init <- ode(State, Time, Bauch.Coupled, Pars.loop)
  # get converged points
  out[[i]] <- ode(init[n,-1], Time, Bauch.Coupled, Pars.loop)[,-1]
}


range.lim <- lapply(out, function(x) apply(x, 2, range))
range.lim <- apply(do.call("rbind", range.lim), 2, range)
plot.variable <- "F_1" # CHOOSE WHICH VARIABLE TO SHOW
plot(0, 0, pch = "", xlab = param.name, ylab = plot.variable,
     xlim = range(param.seq), ylim = range.lim[,plot.variable])
for (i in 1:length(param.seq)) {
  points(rep(param.seq[i], n), out[[i]][,plot.variable])
}





```

```{r sensitivity}

#See: https://search.r-project.org/CRAN/refmans/FME/html/sensFun.html

BauchSensitivity <- function(pars){
  derivs <- function (t, state, pars) {
  
  with (as.list(c(state, pars)), {
  ## now code the model equations
  dF_1dt <-  r_1 * F_1 * (1-F_1) - (h_1 * F_1 * (1-X_1))/(F_1+s_1) - i_2 * F_1 + i_1 * F_2
  dF_2dt <-  r_2 * F_2 * (1-F_2) - (h_2 * F_2 * (1-X_2))/(F_2+s_2) - i_1 * F_2 + i_2 * F_1
  
  dX_1dt <- k_1 * X_1 * (1-X_1) *(1/(F_1 + c_1) - w_1 + d_1*(2*X_1 - 1) + rho_1 * (2*X_2 - 1))
  dX_2dt <- k_2 * X_2 * (1-X_2) *(1/(F_2 + c_2) - w_2 + d_2*(2*X_2 - 1) + rho_2 * (2*X_1 - 1))
  
      return(list(c(dF_1dt, dF_2dt, dX_1dt, dX_2dt)))
    })
  }
  state   <- starting_defaultparams
  tout    <- timespan
  ## ode solves the model by integration ...
  return(as.data.frame(ode(y = state, times = tout, func = derivs,
    parms = pars)))
}


SnsBact <- sensFun(func = BauchSensitivity, parms = params_default, sensvar = "F_1", varscale = 1)

summary(SnsBact)
plot(summary(SnsBact))

plot(SnsBact, type = "b", pch = 15:19, col = 2:6, 
     main = "Sensitivity all vars")

SF <- SnsBact <- sensFun(func = BauchSensitivity, parms = params_default, sensvar = c("F_1", "F_2", "X_1", "X_2"), varscale = 1)

#head(SF)
#tail(SF)

summary(SF, var = TRUE)

plot(SF)
plot(SF, which = c("F_1", "F_2", "X_1", "X_2"))
pm <- par(mfrow = c(1,3))
plot(SF, which = c("F_1", "F_2", "X_1", "X_2"), mfrow = NULL)
plot(SF, mfrow = NULL)
par(mfrow = pm)

## Bivariate sensitivity
pairs(SF)  # same color
pairs(SF, which = "F_1", col = "green", pch = 15)
pairs(SF, which = c("F_1", "F_2"), col = c("green", "blue"))
mtext(outer = TRUE, side = 3, line = -2,
      "Sensitivity functions", cex = 1.5)


## pairwise correlation


SnsBactcorr <- sensFun(func = BauchSensitivity, parms = params_default, sensvar = c("F_1","F_2"), varscale = 1)
correlations <- matrix(cor(SnsBactcorr[,-(1:2)]), 18, 18)
colnames(correlations) <- param_names

correlationplot <- cbind(param1 = data.frame(param_names), correlations) %>%
  pivot_longer(c("r_1", "s_1", "h_1", "k_1", "w_1", "c_1", "d_1", "i_1", "rho_1",
                 "r_2", "s_2", "h_2", "k_2", "w_2", "c_2", "d_2", "i_2", "rho_2"))

ggplot(correlationplot, aes(param_names, name, fill = value)) + geom_tile(width=1) +
  scale_fill_gradient2(
  low = "red",
  mid = "white",
  high = "green",
  #limits = c(1, max(growthRate$lambda_new)),
  midpoint = 0,
  space = "Lab",
  na.value = "grey50") #+
  # scale_x_continuous(name = "Length of closure (months)", expand=c(0,0), breaks = c(1, 3, 6, 9, 12)) +
  # scale_y_continuous(name = "Percent change in survival", expand=c(0,0)) +
  # labs(fill = "Eigenvalue")

```

\newpage

```{r socparamtable, results = 'asis', echo = FALSE}

params_social <- params_default

params_social$r_1 <- 0.4
params_social$r_2 <- 0.35
params_social$w_1 <- 0.2
params_social$w_2 <- 0.35
params_social$h_1 <- 0.25
params_social$h_2 <- 0.5

# params_social$c_1 <- 0.25
# params_social$c_2 <- 0.25


giveparams <- data.frame(matrix(, nrow = 9, ncol = 4))
colnames(giveparams) <- c("Parameter", "Population_1", "Population_2", "Def")

giveparams$Parameter <- c("r", "s", "h", "k", "$\\omega$", "c", "d", "i", "$\\rho$")
giveparams$Population_1 <- c(params_social[1:9])
giveparams$Population_2 <- c(params_social[10:18])

giveparams$Def <- c("Fish net growth",
                    "Supply and demand",
                    "Harvesting efficiency",
                    "Social learning rate",
                    "Conservation cost",
                    "Rarity valuation",
                    "Social norm strength (within pop)",
                    "Fish immigration (from patch)",
                    "Social norm strength (opposite pop)")

knitr::kable(giveparams, format = "pandoc", booktabs = TRUE, caption = "Default parameter values used in this analysis")

```

# Scenarios

## One groups unsustainable practices can tank whole fishery;

params_social$r_1 <- 0.4

params_social$r_2 <- 0.4

params_social$w_1 <- 0.2

params_social$w_2 <- 0.2

params_social$h_1 <- 0.25

params_social$h_2 <- 0.25
Goes up

Vs.

params_social$r_1 <- 0.4

params_social$r_2 <- 0.35

params_social$w_1 <- 0.2

params_social$w_2 <- 0.35

params_social$h_1 <- 0.25

params_social$h_2 <- 0.5
Goes down

BUT increasing rarity val can save that

params_social$c_1 <- 0.25

params_social$c_2 <- 0.25
Goes up
 

```{r socialscenario, fig.cap = "Social inequity scenario where rho_1 = 0.25 and h_1 = 1"}

ode(
  func=Bauch.Coupled,
  y=xstart,
  times=times,
  parms=params_social
) %>%
  as.data.frame() -> out_social

out_social %>%
  gather(variable,value,-time) %>%
  ggplot(aes(x=time,y=value,color=variable))+
  geom_line(size=1)+
  theme_classic()+
  labs(x='time (yr)',y='pop')

```

\newpage

 * Bifurcation plots
 * Sensitivity Analysis: d, rho, s, w, c, i
 * Scenarios
 * Stability and dynamics - Sigdel 2017 good example
 * Go through bauch, thampi, and other papers to see what analysis ideas I can also apply
 * Bauch 2012 does probabilistic sensitivity analysis (PSA). Also see supporting info
 * Phase plane analysis in cressman 2004 and go through your own phase planes to find interesting dynamics
 * Maybe try this: https://cran.r-project.org/web/packages/rootSolve/rootSolve.pdf
