---
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
  pdf_document:
    toc: no
  bookdown::html_document2:
    fig_caption: yes
    toc: no
  word_document: default
bibliography:
- Ch2_citations.bib
- R_Citations.bib
header-includes: \usepackage{setspace}\doublespacing \usepackage{lineno} \usepackage{placeins}
---

\setcounter{page}{57}

\begin{center}
Supplementary Material for:
	
\textbf{\Large Social-ecological models with social hierarchy and space applied to small scale fisheries}
	
\textsc{Sophie Wulfing$^{1*}$ and Easton R. White$^{1}$\\}
\vspace{3 mm}
\normalsize{\indent $^1$Department of Biological Sciences, University of New Hampshire, 03824, NH, USA\\}
$\text{*}$ Corresponding authors: Sophie Wulfing (SophieWulfing@gmail.com) and Dr. Easton White (Easton.White@unh.edu)
\end{center}

\pagebreak


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev="cairo_pdf", cache = TRUE)

library(deSolve)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(kableExtra)
library(FME)
library(ggpubr)

setwd("C:/Users/sophi/Documents/GitHub/UNH_Docs/Proposal/Soc_Ecol_Model")

```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
\linenumbers

Equations 2), 4), and 5) are as follows:

\begin{equation}
\tag{2}
\frac{dX_i}{dt} = k_iX_i(1-X_i)[U_{A,i} - U_{B,i}]
\end{equation}

\begin{equation}
\tag{4}
U_{A,i} = \frac{1}{(F_i + c_i)} + d_iX_i + \rho_i X_j
\end{equation}

\begin{equation}
\tag{5}
U_{B,i} = \omega_i + d_i(1-X_i) + \rho_i(1-X_j)
\end{equation}

Substituting equations 4) and 5) into equation 2 Gives:

$\frac{dX_i}{dt} = k_iX_i(1-X_i)[\frac{1}{(F_i + c_i)} + d_iX_i + \rho_i X_j - \omega_i - d_i(1-X_i) - \rho_i(1-X_j)]$

$\frac{dX_i}{dt} = k_iX_i(1-X_i)[\frac{1}{(F_i + c_i)} - \omega_i + d_i(X_i-1+X_i) + \rho_i(X_j-1+X_j)]$

$\frac{dX_i}{dt} =  k_iX_i(1-X_i) [\frac{1}{F_i+c_i} - \omega_i + d_i(2X_i - 1) + \rho_i(2X_j - 1)]$


